New test of TSO, RT primer and RNA amounts
==========================================

In this design, the TSOs are randomised as much as possible (taking into
account that all do not have the same stock concentrations).

Each 384-well plate is divided in 6 "sextants", interleaved for easier
collection of the reaction products.  Each sextant has a different RNA amount,
and is indexed and amplified independently.  There are 9 × 7 combinations of TSO
and RT primer concentrations, plus one negative control with no RNA.  Each
reaction is assigned to a random well within the sextant.

For each plate replicate, a different random seed is used, to generate
different randomisations.

## Load scripts and libraries.

```{r}
library("magrittr")   # For its diamond pipe operators
library("ggplot2")    # For plotting
library("platetools") # For defining and manipulating 384-well plates
library("plyr")
library("tibble")     # For comfortable command-line operations

# The set_block function that will become part of platetools
source("https://raw.githubusercontent.com/charles-plessy/platetools/76213accb1704d11a2d96fb1f6284d0b46117778/R/set_block.R")

# Do not load the whole smallCAGEqc package, just get the barcode data.
data("nanoCAGE2017barcodes", package = "smallCAGEqc")
```


Selected TSOs
=============

TSOs are transferred as 1, 2 or 4 drops from 3 different starting
concentrations in the source plate.  Let's call each starting concentration
"Tier_A", "Tier_B" or "Tier_C".  Given constraints on remaining volume, some
TSOs can only be Tier_C, etc.  Let's refer to the TSOs by their numerical
identifier.  Tier_N is a set of barcodes that can be used for the negative
control.  See the file [Labcyte-RT6_TSO_stock_to_source.xls](Labcyte-RT6_TSO_stock_to_source.xls)
for details on the cherrypicking of the TSOs.

```{r tiers}
Tier_A <- c(  3, 15, 27, 39, 51, 63, 75
           , 10, 22, 34, 46, 58, 70, 82
           , 11, 23, 35, 47, 59, 71, 83)

Tier_B <- c(  1, 13, 25, 37, 49, 61, 73
           ,  2, 14, 26, 38, 50, 62, 74
           ,  7, 19, 31, 43, 55, 67, 79)

Tier_C <- c(  4, 16, 28, 40, 52, 64, 76
           ,  5, 17, 29, 41, 65, 77, 89
           ,  6, 18, 30, 42, 66, 78, 90)

Tier_N <- c(  8, 20, 32, 44, 68, 80, 92)
```

Set Random seed
===============

```{r setRandomSeed}
set.seed(newseed)
```

Creation of sextants
====================

```{r sextants, dependson=c("tiers", "setRandomSeed")}
createRandomisedSextant <- function(index, rna) {
  df <- data.frame ( BARCODE_ID = c( sample(Tier_A, 21)
                                   , sample(Tier_B, 21)
                                   , sample(Tier_C, 21)
                                   , sample(Tier_N,  1)))
  tsoMaxConc <- 800
  df$TSO_source <- c( rep(tsoMaxConc /  1, 21)
                    , rep(tsoMaxConc /  8, 21)
                    , rep(tsoMaxConc /  64, 21)
                    , tsoMaxConc / 32)
  df$TSO_vol <- c(rep(c(rep(100,7), rep(50,7), rep(25,7)),3), 25)
  df$TSO <- df$TSO_source * df$TSO_vol / 500
  df$RT_PRIMERS <- c(rep(c(0, 1, 2, 4, 8, 16, 24), 9), 1)
  df$RT_PRIMERS_vol <- 25
  df[df$RT_PRIMERS == 0, "RT_PRIMERS_vol"] <- 0
  df$MASTER_MIX_vol <- 350
  df$INDEX <- index
  df$RNA <- rna
  df$RNA_vol <- 25
  df[64, "RNA_vol"] <- 0
  df$H2O_vol <- 500 - df$RNA_vol - df$RT_PRIMERS_vol - df$TSO_vol - df$MASTER_MIX_vol
  df[sample(nrow(df)),]
}
```


Sextants
========

Plate layout
------------

Create a data frame representing the contents of each well.

```{r make_plate_df}
plate           <- tibble(well = platetools::num_to_well(1:384, plate = "384"))
plate$row       <- sub("..$", "", plate$well) %>% factor
plate$col       <- sub("^.",  "", plate$well) %>% as.numeric %>% factor
```

Fill sextants
-------------

```{r fill_sextants, dependson=c("sextants", "make_plate_df")}
plate$sxt <- paste0 (
  c(rep("A", 8), rep("B", 8), rep("C", 8)),
  ifelse (plate$row %in% LETTERS[1:8 * 2], 2, 1)
) %>% factor

plate %<>% arrange(sxt)

plate %<>% cbind( rbind( createRandomisedSextant(indexes[1], 1e5)
                       , createRandomisedSextant(indexes[2], 1e4)
                       , createRandomisedSextant(indexes[3], 1e3)
                       , createRandomisedSextant(indexes[4], 1e2)
                       , createRandomisedSextant(indexes[5], 1e1)
                       , createRandomisedSextant(indexes[6], 1e0))
      ) %>% as.tibble()

plate$BARCODE_ID
plate$INDEX %<>% factor
plate$BARCODE_SEQ <- nanoCAGE2017barcodes[plate$BARCODE_ID, "barcodes"]

plate$RNA_level <- plate$RNA %>% factor(labels = paste0("RNA_", LETTERS[1:6]))
plate$RTP_level <- plate$RT_PRIMERS %>% factor(labels = paste0("RTP_", LETTERS[1:7]))

plate$RNA[plate$RNA_vol == 0] <- 0
plate$plateID <- plateID
```

Summary
=======

```{r plate_summary, dependson="fill_sextants"}
plate 
summary(plate)
write.table(plate, paste0("plate6", variant, ".txt"), sep = "\t", quote = FALSE, row.names = FALSE)
```

Plate maps
==========

```{r}
plateMap <- function(x, title) {
  platetools::raw_map(plate[[x]], well=plate$well, plate="384") +
  ggtitle(title) +
  viridis::scale_fill_viridis(breaks = unique(plate[[x]]))
}

plateMapLog <- function(x, title) {
  platetools::raw_map(plate[[x]], well=plate$well, plate="384") +
  ggtitle(title) +
  viridis::scale_fill_viridis(breaks = unique(plate[[x]]), trans = "log")
}
```

TSO
---

```{r plot_TSO_concentration}
(plot_TSO <- plateMapLog("TSO", "TSO concentration"))
```

```{r plot_TSO_volume}
(plot_TSO_vol <- plateMap("TSO_vol", "TSO volume"))
```

```{r plot_barcode_IDs}
(plot_BARCODES <- platetools::raw_map(plate[["BARCODE_ID"]], well=plate$well, plate="384") +
  ggtitle("Barcode ID")) + theme(legend.position="none")
```

Indexes
-------

```{r plot_Indexes}
(plot_indexes <- platetools::raw_map(plate[["INDEX"]], well=plate$well, plate="384") +
  ggtitle("INDEXes"))
```

H2O
---

```{r plot_H2O_vol}
(plot_H2O_vol <- plateMap("H2O_vol", "H2O volume"))
```

RT primers
----------

```{r plot_RT_concentration}
(plot_RT <- plateMapLog("RT_PRIMERS", "RT primer concentration"))
```

```{r plot_RT_volume}
(plot_RT_vol <- plateMap("RT_PRIMERS_vol", "RT primer volume"))
```

RNA mass
--------

```{r plot_RNA_mass}
(plot_RNA <- plateMapLog("RNA", "RNA mass (ng)"))
```

```{r plot_RNA_vol}
(plot_RNA_vol <- plateMap("RNA_vol", "RNA volume"))
```


Transfer file
=============

## Source plate layout

### TSO

```{r sourcePlate_TSO}
nanoCAGE2017barcodes$well <- unlist(lapply(LETTERS[1:8], function(x) paste(x, sprintf("%02d", 1:12), sep = "")))

tso_block <- data.frame( id  = c(Tier_A, Tier_B, Tier_C, Tier_N)
                       , row = unlist(lapply(LETTERS[1:10], rep, 7))
                       , col = rep(1:7, 10))

tso_block$well <- paste0(tso_block$row, "0", tso_block$col)

tso_block$barcodes <- nanoCAGE2017barcodes[tso_block$id, "barcodes"]

source <- tibble(well = platetools::num_to_well(1:384, plate = "384"))

for (n in 1:nrow(tso_block))
  source %<>% set_block( tso_block[n,"well"]
                       , tso_block[n,"barcodes"]
                       , 20000)
```

### Water

```{r sourcePlate_H2O}
source %<>% set_block("M01~M06", "H2O", 20000)
```

### RNA

```{r sourcePlate_RNA}
levels(plate$RNA %>% factor)
levels(plate$RNA_level)
source %<>% set_block("L01", "RNA_F", 20000)
source %<>% set_block("L02", "RNA_E", 20000)
source %<>% set_block("L03", "RNA_D", 20000)
source %<>% set_block("L04", "RNA_C", 20000)
source %<>% set_block("L05", "RNA_B", 20000)
source %<>% set_block("L06", "RNA_A", 20000)
```

### RNA

```{r sourcePlate_RTP}
levels(plate$RT_PRIMERS %>% factor)
levels(plate$RTP_level)
source %<>% set_block("K01", "RTP_A", 20000)
source %<>% set_block("K02", "RTP_B", 20000)
source %<>% set_block("K03", "RTP_C", 20000)
source %<>% set_block("K04", "RTP_D", 20000)
source %<>% set_block("K05", "RTP_E", 20000)
source %<>% set_block("K06", "RTP_F", 20000)
source %<>% set_block("K07", "RTP_G", 20000)
```

## Plan plate

```{r planPlate}
planPlate <- tibble(well = platetools::num_to_well(1:384, plate = "384"))

for (n in 1:nrow(plate)) {
  planPlate %<>% set_block( plate[[n,"well"]]
                          , plate[[n,"BARCODE_SEQ"]] %>% as.character
                          , plate[[n,"TSO_vol"]])
  planPlate %<>% set_block( plate[[n,"well"]]
                          , "H2O"
                          , plate[[n,"H2O_vol"]])
  planPlate %<>% set_block( plate[[n,"well"]]
                          , plate[[n,"RTP_level"]] %>% as.character
                          , plate[[n,"RT_PRIMERS_vol"]])
  planPlate %<>% set_block( plate[[n,"well"]]
                          , plate[[n,"RNA_level"]] %>% as.character
                          , plate[[n,"RNA_vol"]])
}
planPlate %<>% replace( . == 0, NA)
```

## Plan transfer

```{r transfers}
source("plates.R")
source("echo.R")

planPlate   <- Plate(plate = planPlate)
sourcePlate <- Plate(plate = source)
destPlate   <- Plate(plate = tibble::tibble(well = num_to_well(1:384, plate = "384")))

echo <- planTransfers(sourcePlate, destPlate, planPlate)
echo

transfers <- as.data.frame(echo %>% showLogs)
transfers <- transfers[order(transfers$from),]

colnames(transfers) <- c("Source well", "Destination well", "Transfer volume", "what")

transfers

write.table( transfers[,1:3], file = paste0("Exp7_plate1", variant, ".csv")
           , quote = FALSE, row.names = FALSE, col.names = TRUE, sep = ",")
```

## Session information

```{r sessionInfo}
sessionInfo()
```