---
title: "RT optimisation with the Labcyte Echo 525 (4th plate design)"
output: 
  html_document: 
    fig_height: 5.25
    fig_width: 5.25
    keep_md: yes
    toc: yes
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 100)
```

Differences with the previous design
====================================

In this plate design, all the concentrations are fixed; the only difference
is the TSO barcode.  The goal is to assess differences in efficiency of each
TSO.  This difference may be caused either by the barcode sequence or by
individual variatons of quality between the oligonucleotides during synthesis,
resuspension, etc.

## Load scripts and libraries.

```{r}
library("magrittr")   # For its diamond pipe operators
library("ggplot2")    # For plotting
library("platetools") # For defining and manipulating 384-well plates
library("plyr")
library("tibble")     # For comfortable command-line operations

# The set_block function that will become part of platetools
source("https://raw.githubusercontent.com/charles-plessy/platetools/76213accb1704d11a2d96fb1f6284d0b46117778/R/set_block.R")

# Do not load the whole smallCAGEqc package, just get the barcode data.
data("nanoCAGE2017barcodes", package = "smallCAGEqc")
```


Plate layout
============

Create a data frame representing the contents of each well.

```{r make_plate_df}
plate           <- tibble(well = platetools::num_to_well(1:384, plate = "384"))
plate$row       <- sub("..$", "", plate$well) %>% factor
plate$col       <- sub("^.",  "", plate$well) %>% as.numeric %>% factor
```

MASTER_MIX
==========

350 nL of mastermix added to each well

```{r MASTER_MIX_vol}
plate %<>% set_block("A01~P24", "MASTER_MIX_vol", 350)
```


TSO
===

20.0 μM ( +100 nL of 100 μM )

```{r TSO_concentration}
plate %<>% set_block("A01~P24", "TSO", 100.0)
plate %<>% set_block("A01~P24", "TSO_vol", 100.0)
```

Barcode ID
-----------

The 96 barcodes are tested.

```{r BARCODE_ID}
plate %<>%
  set_block("A01~B24", "BARCODE_ID_norandom", 01:12) %>%
  set_block("C01~D24", "BARCODE_ID_norandom", 13:24) %>%
  set_block("E01~F24", "BARCODE_ID_norandom", 25:36) %>%
  set_block("G01~H24", "BARCODE_ID_norandom", 37:48) %>%
  set_block("I01~J24", "BARCODE_ID_norandom", 49:60) %>%
  set_block("K01~L24", "BARCODE_ID_norandom", 61:72) %>%
  set_block("M01~N24", "BARCODE_ID_norandom", 73:84) %>%
  set_block("O01~P24", "BARCODE_ID_norandom", 85:96)
```

Index
-----

Each TSO will be dispensed 4 times, so 4 indexes will be used.  Tentatively,
they are named `indexA`, `indexB`, `indexC`, and `indexD`.

```{r indexes}
plate %<>%
  arrange(BARCODE_ID_norandom) %>%
  set_block("A01~P24", "INDEX", c("indexA", "indexB", "indexC", "indexD")) %>%
  arrange(well)
```

Randomisation within indexes
----------------------------

```{r randomisation}
plate %<>% arrange(INDEX)
set.seed(1)
plate$BARCODE_ID <- c( sample(1:96)
                     , sample(1:96)
                     , sample(1:96)
                     , sample(1:96))
plate %<>% arrange(well)
plate$INDEX %<>% factor
```

Barcode sequences
-----------------

```{r BARCODE_SEQ}
plate$BARCODE_SEQ <- nanoCAGE2017barcodes[plate$BARCODE_ID,1]
plate$BARCODE_SEQ %<>% factor
```


H2O
====

H2O volume
-----------

0, 25, 50, 75 or 100 nL H2O added to complete RT reaction volume up to 500 nL

```{r H2O_vol}
plate %<>% set_block("A01~P24", "H2O_vol", 0)
```


Negative controls
=================

Let's remove each barcode exactly once.  12 barcodes per index.
The seed is different to the one above.  The plate here has 48 controls.
The seed is re-set so that a sister plate can be made, with 48 other
non-overlaping controls.

```{r}
set.seed(10)
neg_ctl <- sample(1:96)
neg_ctl <- tibble(id = neg_ctl, idx = unlist(lapply(paste0("index", LETTERS[1:8]), rep, 12)))
neg_ctl <- neg_ctl[1:48,]

nc <- match(paste(neg_ctl$id, neg_ctl$idx), paste(plate$BARCODE_ID, plate$INDEX))

plate[nc, "TSO_vol"] <- 0
plate[nc, "H2O_vol"] <- 100
```

RT_PRIMERS
===========

4.0 μM (+ 25 nL from source well at 20 μM)
 
```{r RT_PRIMERS}
plate %<>% set_block("A01~P24", "RT_PRIMERS",  4.0)
plate %<>% set_block("A01~P24", "RT_PRIMERS_vol",  25)
```


RNA
====

RNA concentration
------------------

1 ng (+ 25 nL from source well at 40 ng/μL)

```{r RNA}
plate %<>% set_block("A01~P24", "RNA",  1)
plate %<>% set_block("A01~P24", "RNA_vol",  25)
plate %<>% set_block("A01~P24", "RNA_group", "1ng")
```


Total volume
============

```{r total_volume}
plate$total_volume <- ""
plate$total_volume <- rowSums(plate[, c("MASTER_MIX_vol", "TSO_vol", "RT_PRIMERS_vol", "RNA_vol", "H2O_vol")])
```

RATIO TSO/RT_PRIMERS
=====================

Different ratio of TSO/RT_PRIMERS tested

```{r ratio_TSO_RT_PRIMERS}
plate$PRIMERS_RATIO <- ""
plate$PRIMERS_RATIO <- c(plate$TSO/plate$RT_PRIMERS)
plate$PRIMERS_RATIO <- sub("Inf", "no_RT_PRIMERS", plate$PRIMERS_RATIO)
```

Summary
=======

```{r plate_summary}
plate 
summary(plate)
length(which(plate$total_volume != "NA")) 
sum(is.na(plate$total_volume))
count(plate$BARCODE_ID)
plate$PRIMERS_RATIO
write.table(plate, "plate4.txt", sep = "\t", quote = FALSE, row.names = FALSE)
```

Plate maps
==========

```{r}
plateMap <- function(x, title) {
  platetools::raw_map(plate[[x]], well=plate$well, plate="384") +
  ggtitle(title) +
  viridis::scale_fill_viridis(breaks = unique(plate[[x]]))
}

plateMapLog <- function(x, title) {
  platetools::raw_map(plate[[x]], well=plate$well, plate="384") +
  ggtitle(title) +
  viridis::scale_fill_viridis(breaks = unique(plate[[x]]), trans = "log")
}
```

TSO
---

```{r plot_TSO_concentration}
(plot_TSO <- plateMapLog("TSO", "TSO concentration"))
```

```{r plot_TSO_volume}
(plot_TSO_vol <- plateMap("TSO_vol", "TSO volume"))
```

```{r plot_barcode_IDs}
(plot_BARCODES <- plateMap("BARCODE_ID", "Barcode ID"))
```

Indexes
-------

```{r plot_Indexes}
(plot_RNA_group <- platetools::raw_map(plate[["INDEX"]], well=plate$well, plate="384") +
  ggtitle("INDEXes"))
```

H2O
---

```{r plot_H2O_vol}
(plot_H2O_vol <- plateMap("H2O_vol", "H2O volume"))
```

RT primers
----------

```{r plot_RT_concentration}
(plot_RT <- plateMapLog("RT_PRIMERS", "RT primer concentration"))
```

```{r plot_RT_volume}
(plot_RT_vol <- plateMap("RT_PRIMERS_vol", "RT primer volume"))
```

RNA mass
--------

```{r plot_RNA_mass}
(plot_RNA <- plateMapLog("RNA", "RNA mass (ng)"))
```

```{r plot_RNA_vol}
(plot_RNA_vol <- plateMap("RNA_vol", "RNA volume"))
```

```{r plot_RNA_group}
(plot_RNA_group <- platetools::raw_map(plate[["RNA_group"]], well=plate$well, plate="384") +
  ggtitle("RNA group"))
```


Transfer file for TSOs
======================

## Source plate layout

```{r sourcePlate}
nanoCAGE2017barcodes$well <- unlist(lapply(LETTERS[1:8 * 2 -1], function(x) paste(x, sprintf("%02d", 1:12), sep = "")))

source <- tibble(well = platetools::num_to_well(1:384, plate = "384"))

for (n in 1:96)
  source %<>% set_block( nanoCAGE2017barcodes[n,"well"]
                       , nanoCAGE2017barcodes[n,"barcodes"]
                       , 30000)

source %<>% set_block("P24", "H2O", 30000)
```

## Plan plate

```{r planPlate}
planPlate <- tibble(well = platetools::num_to_well(1:384, plate = "384"))

for (n in 1:nrow(plate)) {
  planPlate %<>% set_block( plate[[n,"well"]]
                          , plate[[n,"BARCODE_SEQ"]] %>% as.character
                          , plate[[n,"TSO_vol"]])
  planPlate %<>% set_block( plate[[n,"well"]]
                          , "H2O"
                          , plate[[n,"H2O_vol"]])
}
planPlate %<>% replace( . == 0, NA)
```

## Plan transfer

```{r transfers}
source("plates.R")
source("echo.R")

planPlate   <- Plate(plate = planPlate)
sourcePlate <- Plate(plate = source)
destPlate   <- Plate(plate = tibble::tibble(well = num_to_well(1:384, plate = "384")))

echo <- planTransfers(sourcePlate, destPlate, planPlate)
echo

transfers <- as.data.frame(echo %>% showLogs)
transfers <- transfers[order(transfers$from),]

transfers

write.table( transfers[,1:3], file = "Exp5_plate1_csv"
           , quote = FALSE, row.names = FALSE, col.names = FALSE, sep = ",")
```

## Session information

```{r sessionInfo}
sessionInfo()
```