---
title: "Labcyte-RT Data Analysis (4th experiment)"
output:
  html_document:
    keep_md: yes
    number_sections: yes
    toc: yes
---

```{r echo = FALSE}
knitr::opts_chunk$set(cache  = TRUE)
knitr::opts_knit$set(verbose = TRUE)
ggplot2::theme_set(ggplot2::theme_bw())
```

Experiment 5
============

Here, we randomised TSO positions 8 times, to see if there was a TSO-specific
bias on the reaction yields.  The bias could be caused either by the barcode
sequence or by the synthesis reaction (we have only one synthesis batch, hence
these two factors are conflated).

The reverse-transcriptase was SuperScript III.

Load R packages
===============

```{r load_R_libraries, message = FALSE}
library("CAGEr")
library("ggplot2")
library("magrittr")
library("MultiAssayExperiment")
library("SummarizedExperiment")
```

MOIRAI shortcuts

```{r moirai_shortcuts, dependson = "load_R_libraries"}
MISEQ_RUN      <- "180411_M00528_0351_000000000-BN3BL"
WORKFLOW       <- "OP-WORKFLOW-CAGEscan-short-reads-v2.1.1"
MOIRAI_STAMP   <- "20180412055135"
MOIRAI_PROJ    <- "project/Labcyte" 
MOIRAI_USER    <- "nanoCAGE2" 
ASSEMBLY       <- "mm9"
BASEDIR        <- "/osc-fs_home/scratch/moirai"
MOIRAI_BASE    <- file.path(BASEDIR, MOIRAI_USER)
MOIRAI_RESULTS <- file.path(MOIRAI_BASE, MOIRAI_PROJ, paste(MISEQ_RUN, WORKFLOW, MOIRAI_STAMP, sep = "."))
```

Load CAGE libraries
===================

Load summary statistics from MOIRAI and polish the names
--------------------------------------------------------

```{r MOIRAI, dependson = "moirai_shortcuts"}
ce <- smallCAGEqc::loadMoiraiStats(
  pipeline  = WORKFLOW,
  multiplex = file.path( MOIRAI_BASE, "input", paste0(MISEQ_RUN, ".multiplex.txt")),
  summary   = file.path( MOIRAI_RESULTS, "text", "summary.txt")) %>% DataFrame

ce$inputFiles <- paste0(MOIRAI_RESULTS, "/CAGEscan_fragments/", ce$samplename, ".bed")

# Discard lines for which input files do not exist.
ce <- ce[sapply(ce$inputFiles, file.exists),]

# Discard lines for which input files are empty.
ce <- ce[file.info(ce$inputFiles)$size != 0,]

ce$inputFilesType <- c("bed")
ce$sampleLabels <- as.character(ce$samplename)

# Replace indexes in group names by RNA amounts extracted from sample sheet.
levels(ce$group) <- system("cut -f 6,8 -d , 180411_M00528_0351_000000000-BN3BL.SampleSheet.csv | grep y_ | sort | cut -f2 -d,", intern = TRUE)

ce$repl <- ce$index
levels(ce$repl) <- system("cut -f 6,8 -d , 180411_M00528_0351_000000000-BN3BL.SampleSheet.csv | grep y_ | sort | cut -f2 -d, | cut -f2 -d_", intern = TRUE)
ce$repl %<>% factor(levels = 1:8)

# Define plate IDs
ce$plateID <- ce$repl
levels(ce$plateID) <- c(rep("M", 4), rep("N", 4))

rownames(ce) %<>% paste(ce$plateID, sep = "_")
ce$sampleLabels <- rownames(ce)

ce
```


Load plate design
-----------------

Using plate 4 and 5 design, see [Labcyte-RT4.md](Labcyte-RT4.md) and 
[Labcyte-RT5.md](Labcyte-RT5.md).

```{r PLATE, dependson = "MOIRAI"}
plate4 <- read.table("plate4.txt", sep = "\t", header = TRUE)
plate5 <- read.table("plate5.txt", sep = "\t", header = TRUE)
plate  <- rbind(plate4, plate5)
plate  <- plate[!duplicated(plate),]

ce %<>% cbind(plate[match( paste(ce$barcode, ce$group)
                         , paste(plate$BARCODE_SEQ, plate$RNA_group)), ])
rm(plate, plate4, plate5)
```


Create a CAGEexp object and load expression data
------------------------------------------------

```{r new_CAGEexp, message = F, echo = FALSE, dependson = "PLATE"}
ce <- CAGEexp( colData    = ce
             , genomeName = "BSgenome.Mmusculus.UCSC.mm9")
```

```{r get_CTSS_data, dependson = "new_CAGEexp"}
getCTSS(ce, useMulticore = TRUE)
removeStrandInvaders(ce)
```


Save the CAGEexp file
---------------------

```{r save_myCAGEexp_object, dependson = "get_CTSS_data"}
saveRDS(ce, "Labcyte-RT_Data_Analysis_4.Rds")
```


Quality controls
================

Custom _scopes_ displaying _strand invasion_ artefacts.

```{r custom_plotAnnot_scopes}
msScope_qcSI <- function(libs) {
  libs$Tag_dust     <- libs$extracted   - libs$rdna - libs$spikes - libs$cleaned
  libs$rDNA         <- libs$rdna
  libs$Spikes       <- libs$spikes
  libs$Unmapped     <- libs$cleaned     - libs$mapped
  libs$Non_proper   <- libs$mapped      - libs$properpairs
  libs$Duplicates   <- libs$properpairs - libs$librarySizes - libs$strandInvaders
  libs$Invaders     <- libs$strandInvaders
  libs$Counts       <- libs$librarySizes
  list( libs    = libs
      , columns = c( "Tag_dust", "rDNA", "Spikes", "Unmapped"
                   , "Non_proper", "Duplicates", "Invaders", "Counts")
      , total   = libs$extracted)
}

msScope_counts <- function(libs) {
  libs$Promoter   <- libs$promoter
  libs$Exon       <- libs$exon
  libs$Intron     <- libs$intron
  libs$Intergenic <- libs$librarySizes - libs$promoter - libs$intron - libs$exon
  libs$Invaders   <- libs$strandInvaders
  list( libs    = libs
      , columns = c("Promoter", "Exon", "Intron", "Intergenic", "Invaders")
      , total   = libs$librarySizes + libs$strandInvaders)
}
```

By RNA input
------------

 * Negative controls with no RNA gave much less sequences than the regula
   samples with RNA.
   
 * The fraction of unmapped reads is larger in RNA negative controls. This might
   suggest that they originate from low-quality sequences that are erroneously
   demultiplexed.
   
 * The amount of oligonucleotide artefacts ("tag dust") is much larger in
   plate _K_.
   
 * The fraction of PCR duplicates increases as the number of counts decreases.
   This is expected and is utilised in CAGEscan assembly.
   
 * The number of non-proper alignments is still quite large.

```{r qc_RNA_quantity, dependson = "get_CTSS_data"}
ggpubr::ggarrange(
  plotAnnot( ce, scope = msScope_qcSI, group = "plateID", normalise = FALSE
           , title = "sequence counts"
           , facet="RNA") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ce, scope = msScope_qcSI, group = "plateID", normalise = TRUE
         , title = "normalised to 100 %"
         , facet="RNA") +
  facet_wrap(~facet, ncol = 1),
  legend = "right",
  common.legend = TRUE
)
```


### Removal of the RNA negative controls

To ease data handling, the negative controls with no RNA are removed.

```{r remove_noRNA_ctls, dependson = "load_and_merge"}
ce <- ce[,ce$RNA != 0]
```

```{r qc_RNA_quantity_raw, dependson = "get_CTSS_data"}
plotAnnot( ce, scope = msScope_qcSI, group = "plateID", normalise = FALSE
         , title = "QC control, by ng of input RNA (sequence counts)"
         , facet="RNA") +
  facet_wrap(~facet, ncol = 1)
```


By RT primer quantity
---------------------

Strangely, libraries made with no RT primers have a QC profile that is in
continuation with the other libraries.  While presence of reads in negative
controls might again be explained contaminations, the amount of sequences in
the "no RT primer" samples is a bit high for such an explanation.

```{r qc_rt_primer_quantity, dependson = "remove_noRNA_ctls"}
ggpubr::ggarrange(
  plotAnnot( ce, scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "sequence counts"
           , facet="plateID") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ce, scope = msScope_qcSI, group = "RT_PRIMERS", normalise = TRUE
         , title = "normalised to 100 %"
         , facet="plateID") +
  facet_wrap(~facet, ncol = 1),
  legend = "right",
  common.legend = TRUE
)
```

Larger amounts of TSO coincides with larger amount of reads in the RT primer
negative control

```{r}
 plotAnnot( ce[,ce$RT_PRIMERS == 0], scope = msScope_qcSI, group = "TSO", normalise = FALSE
           , title = "sequence counts in RT primer negative controls."
           , facet="plateID") 
```


### Removal of the primer-negative controls

To ease data handling (for instance when working with primer ratios), the
negative controls with no primers are removed.

```{r remove_noPrimers_ctls, dependson = "remove_noRNA_ctls"}
ce <- ce[,ce$RT_PRIMERS != 0]

```

Lets tidy the primer ratios after the null primer concentrations were removed.

```{r correct_primer_ratios, dependson = "remove_noPrimers_ctls"}
ce$PRIMERS_RATIO %<>% droplevels
levels(ce$PRIMERS_RATIO) %<>% as.numeric %>% signif(2) %>% as.character
ce$PRIMERS_RATIO %<>% factor(levels = levels(ce$PRIMERS_RATIO) %>% gtools::mixedsort())
```


```{r qc_rt_primer_quantity_clean, dependson = "remove_noRNA_ctls"}
ggpubr::ggarrange(
  plotAnnot( ce, scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "sequence counts"
           , facet="plateID") +
    facet_wrap(~facet, nrow = 1),
  plotAnnot( ce, scope = msScope_qcSI, group = "RT_PRIMERS", normalise = TRUE
         , title = "normalised to 100 %"
         , facet="plateID") +
  facet_wrap(~facet, nrow = 1),
  legend = "right",
  common.legend = TRUE
)
```

rRNA is reduced and non-proper alignments increase as the concentration of
RT primers is lowered.


By TSO concentration
--------------------

Replacing the TSOs in the "B" row (plate _L_) solved the problem of low yield,
compared with just moving them to a different row (plate _K_).

```{r TSO, dependson = "get_CTSS_data"}
ggpubr::ggarrange(
  plotAnnot( ce, scope = msScope_qcSI, group = "TSO", normalise = FALSE
           , title = "sequence counts"
           , facet="plateID") +
    facet_wrap(~facet, nrow = 1),
  plotAnnot( ce, scope = msScope_qcSI, group = "TSO", normalise = TRUE
         , title = "normalised to 100 %"
         , facet="plateID") +
  facet_wrap(~facet, nrow = 1),
  legend = "right",
  common.legend = TRUE
)
```


Annotation with GENCODE
-----------------------

Collect Gencode annotations and gene symbols via a local GENCODE file
(mm9 gencode not available in AnnotationHub)

Most pairs align in intergenic regions.  Is it because of the sequence error
combined with very short read 1s ?

```{r annotate_CTSS, dependson = "get_CTSS_data"}
annotateCTSS(ce, rtracklayer::import.gff("/osc-fs_home/scratch/gmtu/annotation/mus_musculus/gencode-M1/gencode.vM1.annotation.gtf.gz"))
```

```{r CTSS_annotation_plot_per_replicate, dependson = "annotate_CTSS"}
ggpubr::ggarrange(
  plotAnnot( ce, scope = msScope_counts, group = "plateID", normalise = FALSE
           , title = "sequence counts"
           , facet="group") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ce, scope = msScope_counts, group = "plateID", normalise = TRUE
         , title = "normalised to 100 %"
         , facet="group") +
  facet_wrap(~facet, ncol = 1),
  legend = "right",
  common.legend = TRUE
)
```


Other analyses
==============

RT_PRIMER concentration, facetted by RNA mass
---------------------------------------------

```{r rt_primer_quantity, fig.height = 10}
ggpubr::ggarrange(
  plotAnnot( ce, scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "sequence counts"
           , facet="RNA") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ce, scope = msScope_qcSI, group = "RT_PRIMERS", normalise = TRUE
         , title = "normalised to 100 %"
         , facet="RNA") +
  facet_wrap(~facet, ncol = 1),
  legend = "right",
  common.legend = TRUE
)
```

TSO concentration, facetted by RNA mass
---------------------------------------------

### Plate K

```{r TSO_quantity_K, fig.height = 10}
ggpubr::ggarrange(
  plotAnnot( ce[,ce$plateID == "K"], scope = msScope_qcSI, group = "TSO", normalise = FALSE
           , title = "sequence counts"
           , facet="RNA") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ce[,ce$plateID == "K"], scope = msScope_qcSI, group = "TSO", normalise = TRUE
         , title = "normalised to 100 %"
         , facet="RNA") +
  facet_wrap(~facet, ncol = 1),
  legend = "right",
  common.legend = TRUE
)
```

### Plate L

```{r TSO_quantity_L, fig.height = 10}
ggpubr::ggarrange(
  plotAnnot( ce[,ce$plateID == "L"], scope = msScope_qcSI, group = "TSO", normalise = FALSE
           , title = "sequence counts"
           , facet="RNA") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ce[,ce$plateID == "L"], scope = msScope_qcSI, group = "TSO", normalise = TRUE
         , title = "normalised to 100 %"
         , facet="RNA") +
  facet_wrap(~facet, ncol = 1),
  legend = "right",
  common.legend = TRUE
)
```



```{r exploration, fig.height=20, fig.width=20}
ceK <- ce[,ce$plateID == "K"]

ceL <- ce[,ce$plateID == "L"]

ggpubr::ggarrange( legend = "right", common.legend = TRUE,
  plotAnnot( ceL[,], scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "sequence counts, plate L"
           , facet="TSO") +
    facet_wrap(~facet, ncol = 1),
plotAnnot( ceK[,], scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "sequence counts, plate K"
           , facet="TSO") +
    facet_wrap(~facet, ncol = 1)
)

ggpubr::ggarrange( legend = "right", common.legend = TRUE,
  plotAnnot( ceL[,ceL$RNA == .01], scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "plate L, .01 ng RNA"
           , facet="TSO") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ceL[,ceL$RNA == .1], scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "plate L, .1 ng RNA"
           , facet="TSO") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ceL[,ceL$RNA == 1], scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "plate L, 1 ng RNA"
           , facet="TSO") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ceL[,ceL$RNA == 10], scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "plate L, 10 ng RNA"
           , facet="TSO") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ceL[,ceL$RNA == 100], scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "plate L, 100 ng RNA"
           , facet="TSO") +
    facet_wrap(~facet, ncol = 1))


ggpubr::ggarrange( legend = "right", common.legend = TRUE,
  plotAnnot( ceL[,ceK$RNA == .01], scope = msScope_qcSI, group = "TSO", normalise = FALSE
           , title = "plate K, .01 ng RNA"
           , facet="RT_PRIMERS") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ceL[,ceK$RNA == .1], scope = msScope_qcSI, group = "TSO", normalise = FALSE
           , title = "plate K, .1 ng RNA"
           , facet="RT_PRIMERS") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ceL[,ceK$RNA == 1], scope = msScope_qcSI, group = "TSO", normalise = FALSE
           , title = "plate K, 1 ng RNA"
           , facet="RT_PRIMERS") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ceL[,ceK$RNA == 10], scope = msScope_qcSI, group = "TSO", normalise = FALSE
           , title = "plate K, 10 ng RNA"
           , facet="RT_PRIMERS") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ceL[,ceK$RNA == 100], scope = msScope_qcSI, group = "TSO", normalise = FALSE
           , title = "plate K, 100 ng RNA"
           , facet="RT_PRIMERS") +
    facet_wrap(~facet, ncol = 1))

ce1ng <- ce[,ce$RNA == 1]

ggpubr::ggarrange( legend = "right", common.legend = TRUE,
  plotAnnot( ce1ng[,ce1ng$plateID == "L"], scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "sequence counts, plate L"
           , facet="TSO") +
    facet_wrap(~facet, ncol = 1),
plotAnnot( ce1ng[,ce1ng$plateID == "K"], scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "sequence counts, plate K"
           , facet="TSO") +
    facet_wrap(~facet, ncol = 1)
)

```

Optima at 20 μM TSO
-------------------

Best RT primer concentration is 4 μM at 20 μM TSO, for any RNA amount.

```{r ce20}
ce20 <- ce[,ce$TSO == 20]
ggpubr::ggarrange(
  plotAnnot( ce20[,ce20$plateID == "K"], scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "sequence counts, plate K"
           , facet="RNA") +
    facet_wrap(~facet, ncol = 1),
  plotAnnot( ce20[,ce20$plateID == "L"], scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
           , title = "sequence counts, plate L"
           , facet="RNA") +
    facet_wrap(~facet, ncol = 1),
  legend = "right",
  common.legend = TRUE
)
```

Session information
===================

```{r}
sessionInfo()
```