---
title: "Labcyte-RT Data Analysis (3nd experiment, plates 1, 2 and 3.)"
output:
  html_document:
    keep_md: yes
    number_sections: yes
    toc: yes
---

```{r echo = FALSE}
knitr::opts_chunk$set(cache  = TRUE)
knitr::opts_knit$set(verbose = TRUE)
ggplot2::theme_set(ggplot2::theme_bw())
```

Data load and QC in R
=====================

```{r load_R_libraries, message = FALSE}
library("CAGEr")
library("ggplot2")
#library("lattice")
library("magrittr")
library("MultiAssayExperiment")
#library("RColorBrewer")
library("SummarizedExperiment")
library("vegan")
```

MOIRAI shortcuts

```{r moirai_shortcuts}
MISEQ_RUN      <- "180215_M00528_0330_000000000-B4GPD_p456"
WORKFLOW       <- "OP-WORKFLOW-CAGEscan-short-reads-v2.1~rc1"
MOIRAI_STAMP   <- "20180219140315"
MOIRAI_PROJ    <- "project/Labcyte" 
MOIRAI_USER    <- "nanoCAGE2" 
ASSEMBLY       <- "mm9"
BASEDIR        <- "/osc-fs_home/scratch/moirai"
MOIRAI_BASE    <- file.path(BASEDIR, MOIRAI_USER)
MOIRAI_RESULTS <- file.path(MOIRAI_BASE, MOIRAI_PROJ, paste(MISEQ_RUN, WORKFLOW, MOIRAI_STAMP, sep = "."))
```

Load CAGE libraries
===================

Load summary statistics from MOIRAI and polish the names
--------------------------------------------------------

```{r MOIRAI}
ce <- smallCAGEqc::loadMoiraiStats(
  pipeline  = WORKFLOW,
  multiplex = file.path( MOIRAI_BASE, "input", paste0(MISEQ_RUN, ".multiplex.txt")),
  summary   = file.path( MOIRAI_RESULTS, "text", "summary.txt")) %>% DataFrame

ce$inputFiles <- paste0(MOIRAI_RESULTS, "/CAGEscan_fragments/", ce$samplename, ".bed")

# Discard lines for which input files do not exist.
ce <- ce[sapply(ce$inputFiles, file.exists),]

# Discard lines for which input files are empty.
ce <- ce[file.info(ce$inputFiles)$size != 0,]

ce$inputFilesType <- c("bed")
ce$sampleLabels <- as.character(ce$samplename)
ce

# Replace indexes in group names by RNA amounts extracted from sample sheet.
levels(ce$group) <- system("cut -f 6,8 -d , 180123_M00528_0325_000000000-B4PCK.SampleSheet.csv | grep g_ | sort | cut -f2 -d, | cut -f1 -d_", intern = TRUE)

# Sort the levels by RNA amount
ce$group %<>% factor(levels = c("100ng", "10ng", "1ng", "100pg", "10pg"))

ce$repl <- ce$index
levels(ce$repl) <- system("cut -f 6,8 -d , 180123_M00528_0325_000000000-B4PCK.SampleSheet.csv | grep g_ | sort | cut -f 2 -d _ | sed 's/\r//'", intern = TRUE)
ce$repl %<>% factor(levels = 1:3)
```

```{r PLATE}
plate <- read.table("plate.txt", sep = "\t", header = TRUE)
ce %<>% cbind(plate[match( paste(ce$barcode, ce$group)
                         , paste(plate$BARCODE_SEQ, plate$RNA_group)), ])
rm(plate)
```


Create a CAGEexp object and load expression data
------------------------------------------------

```{r new_CAGEexp, message = F, echo = FALSE}
ce <- CAGEexp( colData    = ce
             , genomeName = "BSgenome.Mmusculus.UCSC.mm9")
```

```{r get_CTSS_data}
getCTSS(ce, useMulticore = TRUE)
removeStrandInvaders(ce)
```

Quality controls
----------------

Custom _scopes_ displaying _strand invasion_ artefacts.

```{r custom_plotAnnot_scopes}
msScope_qcSI <- function(libs) {
  libs$Tag_dust     <- libs$extracted   - libs$rdna - libs$spikes - libs$cleaned
  libs$rDNA         <- libs$rdna
  libs$Spikes       <- libs$spikes
  libs$Unmapped     <- libs$cleaned     - libs$mapped
  libs$Non_proper   <- libs$mapped      - libs$properpairs
  libs$Duplicates   <- libs$properpairs - libs$librarySizes - libs$strandInvaders
  libs$Invaders     <- libs$strandInvaders
  libs$Counts       <- libs$librarySizes
  list( libs    = libs
      , columns = c( "Tag_dust", "rDNA", "Spikes", "Unmapped"
                   , "Non_proper", "Duplicates", "Invaders", "Counts")
      , total   = libs$extracted)
}

msScope_counts <- function(libs) {
  libs$Promoter   <- libs$promoter
  libs$Exon       <- libs$exon
  libs$Intron     <- libs$intron
  libs$Intergenic <- libs$librarySizes - libs$promoter - libs$intron - libs$exon
  libs$Invaders   <- libs$strandInvaders
  list( libs    = libs
      , columns = c("Promoter", "Exon", "Intron", "Intergenic", "Invaders")
      , total   = libs$librarySizes + libs$strandInvaders)
}

```

### By RNA input

Negative controls with no RNA gave much less sequences than the regular
samples with RNA.

```{r qc_RNA_quantity_raw}
plotAnnot( ce, scope = msScope_qcSI, group = "RNA", normalise = FALSE
         , title = "QC control, by ng of input RNA (sequence counts)")
```

The normalised QC profile is not much different, therefore
the sequences might be _in silico_ or _in vitro_ contaminations.

```{r qc_RNA_quantity_norm}
plotAnnot( ce, scope = msScope_qcSI, group = "RNA", normalise = TRUE
         , title = "QC control, by ng of input RNA (normalised to 100 %)")
```

### Removal of the RNA negative controls

To ease data handling, the negative controls with no RNA are removed.

```{r}
ce <- ce[,ce$RNA != 0]
```

### By RT primer quantity

Strangely, libraries made with no RT primers have a QC profile that is not
dramatically different from other libraries.  This might again be explained
by contaminations, although the amount of sequences in the "no RT primer"
samples is a bit high for such an explanation.

```{r qc_rt_primer_quantity_raw}
plotAnnot( ce, scope = msScope_qcSI, group = "RT_PRIMERS", normalise = FALSE
         , title = "QC control, by amount of RT primers (in μM)")
```

```{r qc_rt_primer_quantity_norm}
plotAnnot( ce, scope = msScope_qcSI, group = "RT_PRIMERS", normalise = TRUE
         , title = "QC control, by amount of RT primers (normalised to 100 %)")
```

### Removal of the primer-negative controls

To ease data handling (for instance when working with primer ratios), the
negative controls with no primers are removed.

```{r}
ce <- ce[,ce$RT_PRIMERS != 0]
ce$PRIMERS_RATIO %<>% droplevels
ce$PRIMERS_RATIO %<>%
  factor(levels = c( "0.078125", "0.15625", "0.3125", "0.625", "1.25", "2.5"
                   , "5", "10", "20", "40", "80", "160", "320", "640"))
```

### By group and replicates

In all conditions, there are a large number of unmapped sequences.
Close inspection of the alignment files and pipeline reports suggest that
sequence quality was so low that a large number of reads 1 in the pairs have too
many sequence errors to be mapped.

```{r}
plotAnnot(ce, scope = msScope_qcSI, group = "repl", facet="group") +
  facet_wrap(~facet, nrow = 5)
plotAnnot(ce, scope = msScope_qcSI, group = "repl", normalise = FALSE, facet="group") +
  facet_wrap(~facet, nrow = 5)
```

### By TSO concentration

There is a trend decreasing sequence yield with decreasing TSO amounts.
Interestingly, the optimum seems to be lower when RNA quantity is lower.

But what is wrong with TSO == 40 ??  (and to some extent with TSO == 1.25)
We are checking this by making a new experiment where the source plate was
re-prepared from scratch and concentration range has more points near 40.

rRNA rate seems to increase when TSO concentration is reduced.

Strand invation seems to depend more on RNA quantity than on TSO concentration ?!

```{r}
plotAnnot(ce, scope = msScope_qcSI, group = "TSO", normalise = FALSE, facet = "group")
plotAnnot(ce, scope = msScope_qcSI, group = "TSO", normalise = TRUE,  facet = "group")
plotAnnot(ce, scope = msScope_qcSI, group = "TSO", normalise = FALSE, facet = "RT_PRIMERS")
```

Barcode "CTGTCT" gave large amounts of artefacts.  This is not the case for its
nearest similar barcode "CTGCTC".

Barcodes "ATCAGC" and "CACACG" yielded a large amount of extracted reads.

```{r ggplot_wrapper}
ggplot.CAGEexp <- function(data, ...)
  ggplot(as.data.frame(colData(data)), ...)

ggplot(ce, aes(TSO, RT_PRIMERS)) + scale_y_log10()+ scale_x_log10() + geom_raster(aes(fill = librarySizes)) + facet_wrap(~repl + group, ncol = 5) + viridis::scale_fill_viridis() + ggtitle("molecules")

ggplot(ce, aes(TSO, RT_PRIMERS)) + scale_y_log10()+ scale_x_log10() + geom_raster(aes(fill = extracted)) + facet_wrap(~repl + group, ncol = 5) + viridis::scale_fill_viridis() + ggtitle("extracted")

ggplot(ce, aes(TSO, RT_PRIMERS)) + scale_y_log10()+ scale_x_log10() + geom_raster(aes(fill = tagdust / extracted * 100)) + facet_wrap(~repl + group, ncol = 5) + viridis::scale_fill_viridis() + ggtitle("%tagdust")
```

Reactions that received multiple droplets of TSOs did not fail (here we are averaging
very different TSO concentrations, to the point is just to show that for volumes
of 100 and 50, we have roughly the same amount of data.)

```{r}
plotAnnot(ce, scope = msScope_qcSI, group = "TSO_vol", normalise = FALSE)
```

### By primer ratio

```{r}
plotAnnot(ce, scope = msScope_qcSI, group = "PRIMERS_RATIO", normalise = FALSE, facet = "group")
```


Annotation with GENCODE
-----------------------

Collect Gencode annotations and gene symbols via a local GENCODE file
(mm9 gencode not available in AnnotationHub)

```{r annotate_CTSS}
annotateCTSS(ce, rtracklayer::import.gff("/osc-fs_home/scratch/gmtu/annotation/mus_musculus/gencode-M1/gencode.vM1.annotation.gtf.gz"))
```

```{r CTSS_annotation_plot_per_replicate}
plotAnnot(ce, scope = msScope_counts, group = "repl", facet = "group") +
  facet_wrap("facet", nrow = 5)
plotAnnot(ce, scope = msScope_counts, group = "repl", facet = "group", norm = F) +
  facet_wrap("facet", nrow = 5)
```

```{r CTSS_annotation_plot_per_group}
plotAnnot(ce, scope = msScope_counts, group = "TSO", normalise = FALSE, facet = "group") +
  facet_wrap("facet", ncol = 5)
plotAnnot(ce, scope = msScope_counts, group = "TSO", normalise = TRUE,  facet = "group") +
  facet_wrap("facet", ncol = 5)
```

Count QC with TSO concentration 40 removed...

```{r CTSS_annotation_plot_per_group_no40, fig.height=4, fig.width=15}
ce2 <- ce
ce2 <- ce2[,ce2$TSO != 40]
plotAnnot(ce2, scope = "counts", group = "TSO", normalise = FALSE, facet = "group") +
  facet_wrap("facet", ncol = 5) + ggtitle("")
```


```{r CTSS_annotation_heatmaps_all}
ggplot(ce, aes(TSO, RT_PRIMERS)) + scale_y_log10()+ scale_x_log10() + geom_raster(aes(fill = promoter / librarySizes * 100)) + facet_wrap(~repl + group, ncol = 5) + viridis::scale_fill_viridis() + ggtitle("promoter rate")

ggplot(ce, aes(TSO, RT_PRIMERS)) + scale_y_log10()+ scale_x_log10() + geom_raster(aes(fill = (properpairs - librarySizes) / librarySizes)) + facet_wrap(~repl + group, ncol = 5) + viridis::scale_fill_viridis() + ggtitle("PCR duplicates %")
```

```{r CTSS_annotation_heatmaps_per_group, fig.height=4, fig.width=15}
ggplot(ce, aes(TSO, RT_PRIMERS)) + scale_y_log10()+ scale_x_log10() + geom_raster(aes(fill = promoter / librarySizes * 100)) + facet_wrap(~group, ncol = 5) + viridis::scale_fill_viridis() + ggtitle("promoter rate")

ggplot(ce, aes(TSO, RT_PRIMERS)) + scale_y_log10()+ scale_x_log10() + geom_raster(aes(fill = (properpairs - librarySizes) / librarySizes)) + facet_wrap(~group, ncol = 5) + viridis::scale_fill_viridis() + ggtitle("PCR duplicates %")
```

CTSS ANALYSIS
=============

Number of nanoCAGE tags mapping at CTSS positions in each sample
----------------------------------------------------------------
  
```{r ctss-analysis}
ce$l1 <- sapply(CTSStagCountDF(ce), function(x) sum(decode(x) > 0))
```

Gene expression analysis
========================

Make a gene expression table (not really required now).

```{r CTSStoGenes-CAGEexp}
CTSStoGenes(ce)
```

```{r CTSS_ngenes_heatmaps_all}
ggplot(ce, aes(TSO, RT_PRIMERS)) + scale_y_log10()+ scale_x_log10() + geom_raster(aes(fill = genes)) + facet_wrap(~repl + group, ncol = 5) + viridis::scale_fill_viridis() + ggtitle("Gene discovery")
```

```{r CTSS_ngenes_heatmaps_per_group, fig.height=4, fig.width=15}
ggplot(ce, aes(TSO, RT_PRIMERS)) + scale_y_log10()+ scale_x_log10() + geom_raster(aes(fill = genes)) + facet_wrap(~group, ncol = 5) + viridis::scale_fill_viridis() + ggtitle("Gene discovery")
```

Rarefaction
------------

```{r, message = FALSE, fig.height=4, fig.width=15}
ce$r100l1 <- rarefy(t(CTSStagCountDA(ce)),10)
ce$r100l1[librarySizes(ce) < 10] <- NA

ggplot(ce, aes(TSO, RT_PRIMERS)) + scale_y_log10()+ scale_x_log10() + geom_raster(aes(fill = r100l1)) + facet_wrap(~group, ncol = 5) + viridis::scale_fill_viridis() + ggtitle("Richness (on a scale of 10)")
```


```{r precompute_hanabi_objs}
rar1 <- hanabi(CTSStagCountDF(ce), from = 0)
#rarc <- hanabi(assay(consensusClustersSE(myCAGEexp)) %>% as.data.frame, from = 0)
rarg <- hanabi(assay(GeneExpSE(ce)), from = 0)
#save(rar1, rarg, file="rar.Rda") 
```

### Plot TSS discovery

```{r hanabi-TSS_byrep, dev=c('png')}
hanabiPlot(rar1, ylab='number of TSS detected', xlab='number of unique molecule counts', main=paste("TSS discovery"), group=ce$TSO, legend.pos = "bottomright")
```

### Plot Cluster discovery

```{r hanabi-CTSS_clusters_byrep, dev=c('png')}
#hanabiPlot(rarc, ylab='number of CTSS clusters detected', xlab='number of unique molecule counts', main=paste("Cluster discovery"), #group=myCAGEexp$group, legend.pos = "bottomright")
```

### Plot Gene discovery

```{r hanabi-gene_byrep, dev=c('png')}
hanabiPlot(rarg, ylab='number of genes detected', xlab='number of unique molecule counts', main=paste("Gene discovery"), group=ce$TSO, legend.pos = "bottomright")
```

Save the CAGEexp file.

```{r save_myCAGEexp_object}
saveRDS(ce, "Labcyte-RT_Data_Analysis_3b.Rds")
```

```{r}
sessionInfo()
```