---
title: "7th plate design"
subtitle: "RT optimisation with the Labcyte Echo 525: dNTPs and divalent ions."
output: 
  html_document: 
    fig_height: 5.25
    fig_width: 5.25
    keep_md: yes
    toc: yes
    pandoc_args:
     - "--lua-filter=links-to-markdown.lua"
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 100)
```

Influence of dNTP concentration and divalent ions
=================================================

It has been reported ([Lee et al., 2017](https://pubmed.gov/28327113)) that
increasing dNTP concentration increases tailing and favors template switching.
Here, we will replicate and extend the results, and check if the improvement
(if any) has a positive or negative impact on cap specificity.

Increase of [dNTP] must be compensated by increase of divalent ions, otherwise
too many of them will be be chelated and the reaction will be inhibited.
Indeed, [dNTP] is never higher than [Mg<sup>2+</sup>] + [Mg<sup>2+</sup>] in
transcriptome protocols.  Therefore we will not test those combinations.

In addition, addition of manganese, reported to also enhance template switching,
will be tested.

In this design, the TSOs and ion concentrations are randomised as much as
possible (taking into account that all do not have the same stock
concentrations). A fixed TSO, RT primer and RNA concentration is chosen, based
on the results of [Experiment 7](Labcyte-RT_Data_Analysis_7.md).

Each 384-well plate is divided in 6 "sextants", interleaved for easier
collection of the reaction products.  Each sextant has a different
Mn<sup>2+</sup> amount, and is indexed and amplified independently.
There are 9 × 7 combinations of dNTP and Mg<sup>2+</sup>, plus one negative
control with no RNA.  Each reaction is assigned to a random well within the
sextant.

For each plate replicate, a different random seed is used, to generate
different randomisations.

Compared with the previous design, all replicate randomisations and transfers
are calculated at once, to prevent from running out of reagents.


Load scripts and packages
=========================

```{r load_packages}
library("magrittr")   # For its diamond pipe operators
library("ggplot2")    # For plotting
library("platetools") # For defining and manipulating 384-well plates
library("plyr")
library("tibble")     # For comfortable command-line operations

# The set_block function that will become part of platetools
source("https://raw.githubusercontent.com/charles-plessy/platetools/76213accb1704d11a2d96fb1f6284d0b46117778/R/set_block.R")

# Do not load the whole smallCAGEqc package, just get the barcode data.
data("nanoCAGE2017barcodes", package = "smallCAGEqc")
```


Selected TSOs
=============

TSOs are selected according to remaining volume.  See the file [Labcyte-RT7_TSO_stock_to_source.xls](Labcyte-RT7_TSO_stock_to_source.xls)
for details on the cherrypicking of the TSOs.  This time, the concentrations
are adjusted according to the
[concentration check of the stock plate PO_8268526](TSO_concentration_check2.md).

```{r TSOs}
TSOs <- c(  1, 13, 25, 37, 49, 61, 73
         ,  2, 14, 26, 38, 50, 62, 74
         ,  4, 16, 28, 40, 64, 76, 88
         ,  5, 17, 29, 41, 65, 77, 89
         ,  6, 18, 30, 42, 66, 78, 90
         ,  7, 19, 31, 43, 67, 79, 91
         ,  8, 20, 32, 44, 68, 80, 92
         ,  9, 21, 33, 45, 69, 81, 24
         , 10, 22, 34, 46, 70, 82, 36
         , 11, 23, 35, 47, 59, 83, 48)
```

Combinations of magnesium and dNTP concentrations
=================================================

This goes to high extremes, but Smart-Seq 2 has 1 mM dNTP and
12 mM MgCl<sub>2</sub>...

Master Mix "FS" it the standard SSIII first-strand buffer (375 mM KCl,
15 mM MgCl<sub>2</sub>, 250 mM Tris-HCl pH 8.3), and Master Mix "C" (custom)
is the same, without MgCl<sub>2</sub>.

```{r combinations}
combi <- expand.grid( dNTP = c(0.5, 1:5)
                    , Mg   = c(0:3, 3, 4, 6, 9, 12, 15))
combi$MMix <- c(rep("Custom", 24), rep("SSIII", 36))

plot(combi$dNTP, combi$Mg, xlab = "[dNTP] (mM)", ylab ="[Mg] (mM)", main = "Tested combinations", pch = combi$MMix %>% factor %>% as.numeric + 3
     , sub = "Legend: diamonds = standard buffer, crosses = custom")
```

Conveniently, there are 60 remaining combinations, which leaves space for
4 negative control.


Creation of sextants
====================

```{r sextants, dependson=c("tiers", "setRandomSeed")}
createRandomisedSextant <- function(index, Mn) {
  df                       <- data.frame(BARCODE_ID = sample(TSOs, 64))
  df$TSO_source            <- 200
  df$TSO_vol               <- 25
  df$TSO                   <- df$TSO_source * df$TSO_vol / 500
  df$RT_PRIMERS            <- 1
  df$RNA                   <- 10
  df$RT_PRIMERS_RNA_vol    <- 25
  df$MMix                  <- c(combi$MMix, c("Custom", "Custom", "SSIII", "SSIII"))
  df$MASTER_MIX_vol        <- 300
  df$INDEX                 <- index
  df$Mn                    <- Mn
  df$Mn_vol                <- 25
  df[df$Mn == 0, "Mn_vol"] <- 0
  df$Mg                    <- c(combi$Mg, 3, 3, 0, 0)
  df$Mg_vol                <- 25
  df[df$Mg == 0, "Mg_vol"] <- 0
  df$dNTP                  <- c(combi$dNTP, 0, 0, 0, 0)
  df$dNTP_vol              <- 100
  df[61:64, "dNTP_vol"]    <- 0  # Negative controls

  df$H2O_vol <- 500 - with(df, dNTP_vol + Mn_vol + Mg_vol +
                               TSO_vol + RT_PRIMERS_RNA_vol +
                               MASTER_MIX_vol)
  df[sample(nrow(df)),]
}
```


Create plates
=============

```{r make_plates, dependson=c("sextants")}
mkPlate <- function(randomSeed, plateID, indexes) {
  plate           <- tibble(well = platetools::num_to_well(1:384, plate = "384"))
  plate$row       <- sub("..$", "", plate$well) %>% factor
  plate$col       <- sub("^.",  "", plate$well) %>% as.numeric %>% factor

  set.seed(randomSeed)
  plate$sxt <- paste0 (
    c(rep("A", 8), rep("B", 8), rep("C", 8)),
    ifelse (plate$row %in% LETTERS[1:8 * 2], 2, 1)
  ) %>% factor
  
  plate %<>% arrange(sxt)
  
  plate %<>% cbind( rbind( createRandomisedSextant(indexes[1], 0)
                         , createRandomisedSextant(indexes[2], 1)
                         , createRandomisedSextant(indexes[3], 2)
                         , createRandomisedSextant(indexes[4], 3)
                         , createRandomisedSextant(indexes[5], 4)
                         , createRandomisedSextant(indexes[6], 5))
        ) %>% as.tibble()
  
  plate$INDEX %<>% factor
  plate$BARCODE_SEQ <- nanoCAGE2017barcodes[plate$BARCODE_ID, "barcodes"]
  
  plate$Mg_level     <- plate$Mg   %>% factor(labels = paste0("Mg_",   LETTERS[1:9]))
  plate$Mn_level     <- plate$Mn   %>% factor(labels = paste0("Mn_",   LETTERS[1:6]))
  plate$dNTP_level   <- plate$dNTP %>% factor(labels = paste0("dNTP_", LETTERS[1:7]))
  plate$plateID <- plateID
  plate
}

plate7a <- mkPlate(1, "U", c("TAAGGCGA", "CGTACTAG", "AGGCAGAA", "TCCTGAGC", "GGACTCCT", "TAGGCATG"))
plate7b <- mkPlate(2, "V", c("CTCTCTAC", "CGAGGCTG", "AAGAGGCA", "GTAGAGGA", "GCTCATGA", "ATCTCAGG"))
plate7c <- mkPlate(3, "W", c("ACTCGCTA", "GGAGCTAC", "GCGTAGTA", "CGGAGCCT", "TACGCTGC", "ATGCGCAG"))
plate7d <- mkPlate(4, "X", c("TAGCGCTC", "ACTGAGCG", "CCTAAGAC", "CGATCAGT", "TGCAGCTA", "TCGACGTC"))
```


Summary
=======

```{r plate_summary, dependson="make_plates"}
writePlate <- function(plate) {
  plateName <- deparse(substitute(plate))
  write.table( plate
             , paste0(plateName, ".txt")
             , sep = "\t"
             , quote = FALSE
             , row.names = FALSE)
}

writePlate(plate7a)
writePlate(plate7b)
writePlate(plate7c)
writePlate(plate7d)
```


Plate maps for randomisation a
==============================

```{r}
plateMap <- function(x, title) {
  plate <- plate7a
  platetools::raw_map(plate[[x]], well=plate$well, plate="384") +
  ggtitle(title) +
  viridis::scale_fill_viridis(breaks = unique(plate[[x]]))
}

plateMapLog <- function(x, title) {
  plate <- plate7a
  platetools::raw_map(plate[[x]], well=plate$well, plate="384") +
  ggtitle(title) +
  viridis::scale_fill_viridis(breaks = unique(plate[[x]]), trans = "log")
}
```

dNTP, Mg, Mn
------------

```{r plot_dNTP_concentration}
(plot_dNTP <- plateMapLog("dNTP", "dNTP concentration"))
```

```{r plot_Mg_concentration}
(plot_Mg   <- plateMapLog("Mg",   "Mg concentration"))
```

```{r plot_Mn_concentration}
(plot_Mn   <- plateMapLog("Mn",   "Mn concentration"))
```

Barcode IDs
-----------

```{r plot_barcode_IDs}
(plot_BARCODES <- platetools::raw_map(plate7a[["BARCODE_ID"]], well=plate7a$well, plate="384") +
  ggtitle("Barcode ID")) + theme(legend.position="none")
```

Indexes
-------

```{r plot_Indexes}
(plot_indexes <- platetools::raw_map(plate7a[["INDEX"]], well=plate7a$well, plate="384") +
  ggtitle("INDEXes"))
```

H2O
---

```{r plot_H2O_vol}
(plot_H2O_vol <- plateMap("H2O_vol", "H2O volume"))
```


Transfer file
=============

## Source plate layout

### TSOs in wells A01~J07

Source plate concentration is 200 μM, for a 10 μM final concentration
(25 nL transfer).  This time, the dilution factors calculated from the
concentration checks in the stock plate have been applied.  The TSOs will be in
wells A1 to J7 of the source plate.

```{r sourcePlate_TSO}
tso_block <- data.frame( id  = TSOs
                       , row = unlist(lapply(LETTERS[1:10], rep, 7))
                       , col = rep(1:7, 10))

tso_block$well <- paste0(tso_block$row, "0", tso_block$col)

tso_block$barcodes <- nanoCAGE2017barcodes[tso_block$id, "barcodes"]

source <- tibble(well = platetools::num_to_well(1:384, plate = "384"))

for (n in 1:nrow(tso_block))
  source %<>% set_block( tso_block[n,"well"]
                       , tso_block[n,"barcodes"]
                       , 20000)
```


### RNA / RT primer mix in well K01

This time it was needed to combine RNA and RT primers to reduce the total volume.
Fixed RT primer concentration and RNA amount: 20 μM RT primer ; 400 ng / μL RNA.

```{r sourcePlate_RNA_RTP}
source %<>% set_block("K01", "RT_PRIMERS_RNA", 60000)
```


### Water in well K02

```{r sourcePlate_H2O}
source %<>% set_block("K02", "H2O", 60000)
```


### Mn<sup>2+</sup> in wells L01 to L05

Wells L01 to L05 filled with 20, 40, 60, 80 and 100 μM MnCl<sub>2</sub> respectively.

```{r sourcePlate_Mn}
levels(plate7a$Mn %>% factor)
levels(plate7a$Mn_level)
source %<>% set_block("L01", "Mn_B", 60000)
source %<>% set_block("L02", "Mn_C", 60000)
source %<>% set_block("L03", "Mn_D", 60000)
source %<>% set_block("L04", "Mn_E", 60000)
source %<>% set_block("L05", "Mn_F", 60000)
```


### Mg<sup>2+</sup> in wells M01 to M08

Wells M01 to M08 filled with 20, 40, 60, 80, 120, 180, 240, and 300 μM
MgCl<sub>2</sub> respectively.


```{r sourcePlate_Mg}
levels(plate7a$Mg %>% factor)
levels(plate7a$Mg_level)
source %<>% set_block("M01", "Mg_B", 60000)
source %<>% set_block("M02", "Mg_C", 60000)
source %<>% set_block("M03", "Mg_D", 60000)
source %<>% set_block("M04", "Mg_E", 60000)
source %<>% set_block("M05", "Mg_F", 60000)
source %<>% set_block("M06", "Mg_G", 60000)
source %<>% set_block("M07", "Mg_H", 60000)
source %<>% set_block("M08", "Mg_I", 60000)
```


### dNTPs

Wells N01 to N06 filled with 2.5, 5, 10, 15, 20, 25 mM dNTPs respectively.

```{r sourcePlate_dNTP}
levels(plate7a$dNTP %>% factor)
levels(plate7a$dNTP_level)
source %<>% set_block("N01", "dNTP_B", 60000)
source %<>% set_block("N02", "dNTP_C", 60000)
source %<>% set_block("N03", "dNTP_D", 60000)
source %<>% set_block("N04", "dNTP_E", 60000)
source %<>% set_block("N05", "dNTP_F", 60000)
source %<>% set_block("N06", "dNTP_G", 60000)
```


### Mastermixes

Wells O01 to O06 filled with custom mastermix; wells P01 to P07 filled
with standard mastermix.

```{r sourcePlate_MM}
levels(plate7a$MMix %>% factor)
source %<>% set_block("O01~O06", "Custom", 60000)
source %<>% set_block("P01~P07", "SSIII",  60000)
```


## Plan plate

```{r planPlate}
mkPlanPlate <- function(plate) {
  planPlate <- tibble(well = platetools::num_to_well(1:384, plate = "384"))
  
  for (n in 1:nrow(plate)) {
    planPlate %<>% set_block( plate[[n,"well"]]
                            , plate[[n,"BARCODE_SEQ"]] %>% as.character
                            , plate[[n,"TSO_vol"]])
    
    planPlate %<>% set_block( plate[[n,"well"]]
                            , "H2O"
                            , plate[[n,"H2O_vol"]])
    
    planPlate %<>% set_block( plate[[n,"well"]]
                            , "RT_PRIMERS_RNA"
                            , plate[[n,"RT_PRIMERS_RNA_vol"]])
    
    planPlate %<>% set_block( plate[[n,"well"]]
                            , plate[[n,"Mg_level"]] %>% as.character
                            , plate[[n,"Mg_vol"]])
    
    planPlate %<>% set_block( plate[[n,"well"]]
                            , plate[[n,"Mn_level"]] %>% as.character
                            , plate[[n,"Mn_vol"]])
    
    planPlate %<>% set_block( plate[[n,"well"]]
                            , plate[[n,"dNTP_level"]] %>% as.character
                            , plate[[n,"dNTP_vol"]])
    
    planPlate %<>% set_block( plate[[n,"well"]]
                            , plate[[n,"MMix"]] %>% as.character
                            , plate[[n,"MASTER_MIX_vol"]])
  }
  planPlate %<>% replace( . == 0, NA)
  planPlate <- planPlate[,sapply(planPlate, function(x) !all(is.na(x)))]
  planPlate
}
```

## Plan transfer

```{r transfers_a}
source("plates.R")
source("echo.R")

planPlate7a <- Plate(plate = mkPlanPlate(plate7a))
sourcePlate <- Plate(plate = source, deadVolume = 16000)
destPlate   <- Plate(plate = tibble::tibble(well = num_to_well(1:384, plate = "384")))

echo <- planTransfers(sourcePlate, destPlate, planPlate7a)
echo
echo@source@plate[,-1] %>% colSums(na.rm = TRUE)

transfers7a <- as.data.frame(echo %>% showLogs)
transfers7a <- transfers7a[order(transfers7a$from),]

colnames(transfers7a) <- c("Source well", "Destination well", "Transfer volume", "what")

writeMMixTransfers <- function(transfers, filename) {
  transfers <- transfers[transfers$what %in% c("SSIII","Custom"),]
  write.table( transfers[,1:3], file = filename
             , quote = FALSE, row.names = FALSE, col.names = TRUE, sep = ",")
}

writeOtherTransfers <- function(transfers, filename) {
  transfers <- transfers[!transfers$what %in% c("SSIII","Custom"),]
  write.table( transfers[,1:3], file = filename
             , quote = FALSE, row.names = FALSE, col.names = TRUE, sep = ",")
}

writeMMixTransfers (transfers7a, "Exp8_plate1_mastermix.csv")
writeOtherTransfers(transfers7a, "Exp8_plate1_other-reagents.csv")
```

```{r transfers_b}
planPlate7b <- Plate(plate = mkPlanPlate(plate7b))
sourcePlate <- echo@source

echo <- planTransfers(sourcePlate, destPlate, planPlate7b)
echo
echo@source@plate[,-1] %>% colSums(na.rm = TRUE)

transfers7b <- as.data.frame(echo %>% showLogs)
transfers7b <- transfers7b[order(transfers7b$from),]

writeMMixTransfers (transfers7b, "Exp8_plate2_mastermix.csv")
writeOtherTransfers(transfers7b, "Exp8_plate2_other-reagents.csv")
```

```{r transfers_c}
planPlate7c <- Plate(plate = mkPlanPlate(plate7c))
sourcePlate <- echo@source

echo <- planTransfers(sourcePlate, destPlate, planPlate7c)
echo
echo@source@plate[,-1] %>% colSums(na.rm = TRUE)

transfers7c <- as.data.frame(echo %>% showLogs)
transfers7c <- transfers7c[order(transfers7c$from),]

writeMMixTransfers (transfers7c, "Exp8_plate3_mastermix.csv")
writeOtherTransfers(transfers7c, "Exp8_plate3_other-reagents.csv")
```

```{r transfers_d}
planPlate7d <- Plate(plate = mkPlanPlate(plate7d))
sourcePlate <- echo@source

echo <- planTransfers(sourcePlate, destPlate, planPlate7d)
echo
echo@source@plate[,-1] %>% colSums(na.rm = TRUE)

transfers7d <- as.data.frame(echo %>% showLogs)
transfers7d <- transfers7d[order(transfers7d$from),]

writeMMixTransfers (transfers7d, "Exp8_plate4_mastermix.csv")
writeOtherTransfers(transfers7d, "Exp8_plate4_other-reagents.csv")
```

## Inline copy of the tables

```{r}
plate7a %>% as.data.frame
plate7b %>% as.data.frame
plate7c %>% as.data.frame
plate7d %>% as.data.frame
transfers7a
transfers7b
transfers7c
transfers7d
```

## Session information

```{r sessionInfo}
sessionInfo()
```