---
title: "RT optimisation with the Labcyte Echo 525 (December 2017)"
output: 
  html_document: 
    keep_md: yes
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 120)
```

Load scripts and libraries.

```{r}
# We use the R library called "magrittr", to build the data frame describing the
#384-well plate.
library(magrittr)
library(plyr)
# The "setBlock" function to update values in the data frame is shared by multiple
# Rmarkdown files, so it is stored in a separate source file.
source('setBlock.R')
# The "makePlate384" function outputs a data frame representing a 384-well plate
# (one well per row).
source('makePlate384.R')
```

## Plate layout

Create a data frame representing the contents of each well.

```{r create-plate}
plate <- makePlate384()
```

MASTER_MIX
==========

Master mix volume 
------------------

350 nL of mastermix added to each well

```{r MASTER_MIX_vol}
plate %<>%
  setBlock("A01~P21", "MASTER_MIX_vol", 350) %>%
  setBlock("K22~L24", "MASTER_MIX_vol", 350) %>%
  setBlock("M23", "MASTER_MIX_vol", 350) %>%
  setBlock("N22~O24", "MASTER_MIX_vol", 350) %>%
  setBlock("P23", "MASTER_MIX_vol", 350)
```

TSO
===

TSO concentration
------------------

 - 80 uM (  +100 nL of 400 uM A1-A7 source wells) in A1-A7, A8-A14, A15-A21, K1-K7, N1-N7 
 - 40 uM ( +50 nL of 400 uM A1-A7 source wells) in B1-B7, B8-B14, B15-B21, L1-L7, O1-O7
 - 20 uM ( +25 nL of 400 uM A1-A7 source wells) in C1-C7, C8-C14, C15-C21, M1-M7, P1-P7
 - 10 uM ( +100 nL of 50 uM source wells) in D1-D7, D8-D14, D15-D21, K8-K14, N8-N14 + J1-J7, J8-J14, J15-J21, K22-K24, L22-L24, M23, N22-N24, O22-O24, P23
 - 5 uM ( +50 nL of 50 uM source wells) in E1-E7, E8-E14, E15-E21, L8-L14, O8-O14
 - 2,5 uM (+25 nL of 50 uM source wells) in F1-F7, F8-F14, F15-F21, M8-M14, P8-P14
 - 1,25 uM (+100 nL of 6,25 uM source wells) in G1-G7, G8-G14, G15-G21, K15-K21, N15-N21
 - 0,625 uM (+50 nL of 6,25 uM source wells) in H1-H7, H8-H14, H15-H21, L15-L21, O15-O21
 - 0,3125 uM (+25 nL of 6,25 uM source wells) in I1-I7, I8-I14, I15-I21, M15-M21, P15-P21
 

```{r TSO_concentration}
plate %<>%
  setBlock("A01~A21", "TSO", 80) %>%
  setBlock("K01~K07", "TSO", 80) %>%
  setBlock("N01~N07", "TSO", 80) %>%
  setBlock("B01~B21", "TSO", 40) %>%
  setBlock("L01~L07", "TSO", 40) %>%
  setBlock("O01~O07", "TSO", 40) %>%
  setBlock("C01~C21", "TSO", 20) %>%
  setBlock("M01~M07", "TSO", 20) %>%
  setBlock("P01~P07", "TSO", 20) %>%
  setBlock("D01~D21", "TSO", 10) %>%
  setBlock("K08~K14", "TSO", 10) %>%
  setBlock("N08~N14", "TSO", 10) %>%
  setBlock("J01~J21", "TSO", 10) %>%
  setBlock("K22~K24", "TSO", 10) %>%
  setBlock("L22~L24", "TSO", 10) %>%
  setBlock("M23", "TSO", 10) %>%
  setBlock("N22~N24", "TSO", 10) %>%
  setBlock("O22~O24", "TSO", 10) %>%
  setBlock("P23", "TSO", 10) %>% 
  setBlock("E01~E21", "TSO", 5) %>%
  setBlock("L08~L14", "TSO", 5) %>%
  setBlock("O08~O14", "TSO", 5) %>%
  setBlock("F01~F21", "TSO", 2.5) %>%
  setBlock("M08~M14", "TSO", 2.5) %>%
  setBlock("P08~P14", "TSO", 2.5) %>%
  setBlock("G01~G21", "TSO", 1.25) %>%
  setBlock("K15~K21", "TSO", 1.25) %>%
  setBlock("N15~N21", "TSO", 1.25) %>%
  setBlock("H01~H21", "TSO", 0.625) %>%
  setBlock("L15~L21", "TSO", 0.625) %>%
  setBlock("O15~O21", "TSO", 0.625) %>%
  setBlock("I01~I21", "TSO", 0.3125) %>%
  setBlock("M15~M21", "TSO", 0.3125) %>%
  setBlock("P15~P21", "TSO", 0.3125)
```

TSO volume
-----------

25 (1-drop), 50 (2-drops) or 100 nL (4-drops) of TSO added, depending on the final concentration

```{r TSO_volume}
plate %<>%
  setBlock("A01~A21", "TSO_vol", 100) %>%
  setBlock("K01~K07", "TSO_vol", 100) %>%
  setBlock("N01~N07", "TSO_vol", 100) %>%
  setBlock("B01~B21", "TSO_vol", 50) %>%
  setBlock("L01~L07", "TSO_vol", 50) %>%
  setBlock("O01~O07", "TSO_vol", 50) %>%
  setBlock("C01~C21", "TSO_vol", 25) %>%
  setBlock("M01~M07", "TSO_vol", 25) %>%
  setBlock("P01~P07", "TSO_vol", 25) %>%
  setBlock("D01~D21", "TSO_vol", 100) %>%
  setBlock("K08~K14", "TSO_vol", 100) %>%
  setBlock("N08~N14", "TSO_vol", 100) %>%
  setBlock("J01~J21", "TSO_vol", 100) %>%
  setBlock("K22~K24", "TSO_vol", 100) %>%
  setBlock("L22~L24", "TSO_vol", 100) %>%
  setBlock("M23", "TSO_vol", 100) %>%
  setBlock("N22~N24", "TSO_vol", 100) %>%
  setBlock("O22~O24", "TSO_vol", 100) %>%
  setBlock("P23", "TSO_vol", 100) %>% 
  setBlock("E01~E21", "TSO_vol", 50) %>%
  setBlock("L08~L14", "TSO_vol", 50) %>%
  setBlock("O08~O14", "TSO_vol", 50) %>%
  setBlock("F01~F21", "TSO_vol", 25) %>%
  setBlock("M08~M14", "TSO_vol", 25) %>%
  setBlock("P08~P14", "TSO_vol", 25) %>%
  setBlock("G01~G21", "TSO_vol", 100) %>%
  setBlock("K15~K21", "TSO_vol", 100) %>%
  setBlock("N15~N21", "TSO_vol", 100) %>%
  setBlock("H01~H21", "TSO_vol", 50) %>%
  setBlock("L15~L21", "TSO_vol", 50) %>%
  setBlock("O15~O21", "TSO_vol", 50) %>%
  setBlock("I01~I21", "TSO_vol", 25) %>%
  setBlock("M15~M21", "TSO_vol", 25) %>%
  setBlock("P15~P21", "TSO_vol", 25) 
```

Barcode ID
-----------

70 barcodes used for each RNA concentration tested

```{r BARCODE_ID}
plate %<>%
  setBlock("A01", "BARCODE_ID", 01) %>%
  setBlock("A02", "BARCODE_ID", 02) %>%
  setBlock("A03", "BARCODE_ID", 03) %>%
  setBlock("A04", "BARCODE_ID", 04) %>%
  setBlock("A05", "BARCODE_ID", 05) %>%
  setBlock("A06", "BARCODE_ID", 06) %>%
  setBlock("A07", "BARCODE_ID", 07) %>%  
  setBlock("A08", "BARCODE_ID", 01) %>%
  setBlock("A09", "BARCODE_ID", 02) %>%
  setBlock("A10", "BARCODE_ID", 03) %>%
  setBlock("A11", "BARCODE_ID", 04) %>%
  setBlock("A12", "BARCODE_ID", 05) %>%
  setBlock("A13", "BARCODE_ID", 06) %>%
  setBlock("A14", "BARCODE_ID", 07) %>%  
  setBlock("A15", "BARCODE_ID", 01) %>%
  setBlock("A16", "BARCODE_ID", 02) %>%
  setBlock("A17", "BARCODE_ID", 03) %>%
  setBlock("A18", "BARCODE_ID", 04) %>%
  setBlock("A19", "BARCODE_ID", 05) %>%
  setBlock("A20", "BARCODE_ID", 06) %>%
  setBlock("A21", "BARCODE_ID", 07) %>%  
  setBlock("K01", "BARCODE_ID", 01) %>%
  setBlock("K02", "BARCODE_ID", 02) %>%
  setBlock("K03", "BARCODE_ID", 03) %>%
  setBlock("K04", "BARCODE_ID", 04) %>%
  setBlock("K05", "BARCODE_ID", 05) %>%
  setBlock("K06", "BARCODE_ID", 06) %>%
  setBlock("K07", "BARCODE_ID", 07) %>%    
  setBlock("N01", "BARCODE_ID", 01) %>%
  setBlock("N02", "BARCODE_ID", 02) %>%
  setBlock("N03", "BARCODE_ID", 03) %>%
  setBlock("N04", "BARCODE_ID", 04) %>%
  setBlock("N05", "BARCODE_ID", 05) %>%
  setBlock("N06", "BARCODE_ID", 06) %>%
  setBlock("N07", "BARCODE_ID", 07) %>%  
  setBlock("B01", "BARCODE_ID", 08) %>%
  setBlock("B02", "BARCODE_ID", 09) %>%
  setBlock("B03", "BARCODE_ID", 10) %>%
  setBlock("B04", "BARCODE_ID", 11) %>%
  setBlock("B05", "BARCODE_ID", 12) %>%
  setBlock("B06", "BARCODE_ID", 13) %>%
  setBlock("B07", "BARCODE_ID", 14) %>%  
  setBlock("B08", "BARCODE_ID", 08) %>%
  setBlock("B09", "BARCODE_ID", 09) %>%
  setBlock("B10", "BARCODE_ID", 10) %>%
  setBlock("B11", "BARCODE_ID", 11) %>%
  setBlock("B12", "BARCODE_ID", 12) %>%
  setBlock("B13", "BARCODE_ID", 13) %>%
  setBlock("B14", "BARCODE_ID", 14) %>%  
  setBlock("B15", "BARCODE_ID", 08) %>%
  setBlock("B16", "BARCODE_ID", 09) %>%
  setBlock("B17", "BARCODE_ID", 10) %>%
  setBlock("B18", "BARCODE_ID", 11) %>%
  setBlock("B19", "BARCODE_ID", 12) %>%
  setBlock("B20", "BARCODE_ID", 13) %>%
  setBlock("B21", "BARCODE_ID", 14) %>% 
  setBlock("L01", "BARCODE_ID", 08) %>%
  setBlock("L02", "BARCODE_ID", 09) %>%
  setBlock("L03", "BARCODE_ID", 10) %>%
  setBlock("L04", "BARCODE_ID", 11) %>%
  setBlock("L05", "BARCODE_ID", 12) %>%
  setBlock("L06", "BARCODE_ID", 13) %>%
  setBlock("L07", "BARCODE_ID", 14) %>% 
  setBlock("O01", "BARCODE_ID", 08) %>%
  setBlock("O02", "BARCODE_ID", 09) %>%
  setBlock("O03", "BARCODE_ID", 10) %>%
  setBlock("O04", "BARCODE_ID", 11) %>%
  setBlock("O05", "BARCODE_ID", 12) %>%
  setBlock("O06", "BARCODE_ID", 13) %>%
  setBlock("O07", "BARCODE_ID", 14) %>% 
  setBlock("C01", "BARCODE_ID", 15) %>%
  setBlock("C02", "BARCODE_ID", 16) %>%
  setBlock("C03", "BARCODE_ID", 17) %>%
  setBlock("C04", "BARCODE_ID", 18) %>%
  setBlock("C05", "BARCODE_ID", 19) %>%
  setBlock("C06", "BARCODE_ID", 20) %>%
  setBlock("C07", "BARCODE_ID", 21) %>%   
  setBlock("C08", "BARCODE_ID", 15) %>%
  setBlock("C09", "BARCODE_ID", 16) %>%
  setBlock("C10", "BARCODE_ID", 17) %>%
  setBlock("C11", "BARCODE_ID", 18) %>%
  setBlock("C12", "BARCODE_ID", 19) %>%
  setBlock("C13", "BARCODE_ID", 20) %>%
  setBlock("C14", "BARCODE_ID", 21) %>%  
  setBlock("C15", "BARCODE_ID", 15) %>%
  setBlock("C16", "BARCODE_ID", 16) %>%
  setBlock("C17", "BARCODE_ID", 17) %>%
  setBlock("C18", "BARCODE_ID", 18) %>%
  setBlock("C19", "BARCODE_ID", 19) %>%
  setBlock("C20", "BARCODE_ID", 20) %>%
  setBlock("C21", "BARCODE_ID", 21) %>%
  setBlock("M01", "BARCODE_ID", 15) %>%
  setBlock("M02", "BARCODE_ID", 16) %>%
  setBlock("M03", "BARCODE_ID", 17) %>%
  setBlock("M04", "BARCODE_ID", 18) %>%
  setBlock("M05", "BARCODE_ID", 19) %>%
  setBlock("M06", "BARCODE_ID", 20) %>%
  setBlock("M07", "BARCODE_ID", 21) %>%
  setBlock("P01", "BARCODE_ID", 15) %>%
  setBlock("P02", "BARCODE_ID", 16) %>%
  setBlock("P03", "BARCODE_ID", 17) %>%
  setBlock("P04", "BARCODE_ID", 18) %>%
  setBlock("P05", "BARCODE_ID", 19) %>%
  setBlock("P06", "BARCODE_ID", 20) %>%
  setBlock("P07", "BARCODE_ID", 21) %>%
  setBlock("D01", "BARCODE_ID", 22) %>%
  setBlock("D02", "BARCODE_ID", 23) %>%
  setBlock("D03", "BARCODE_ID", 24) %>%
  setBlock("D04", "BARCODE_ID", 25) %>%
  setBlock("D05", "BARCODE_ID", 26) %>%
  setBlock("D06", "BARCODE_ID", 27) %>%
  setBlock("D07", "BARCODE_ID", 28) %>%   
  setBlock("D08", "BARCODE_ID", 22) %>%
  setBlock("D09", "BARCODE_ID", 23) %>%
  setBlock("D10", "BARCODE_ID", 24) %>%
  setBlock("D11", "BARCODE_ID", 25) %>%
  setBlock("D12", "BARCODE_ID", 26) %>%
  setBlock("D13", "BARCODE_ID", 27) %>%
  setBlock("D14", "BARCODE_ID", 28) %>% 
  setBlock("D15", "BARCODE_ID", 22) %>%
  setBlock("D16", "BARCODE_ID", 23) %>%
  setBlock("D17", "BARCODE_ID", 24) %>%
  setBlock("D18", "BARCODE_ID", 25) %>%
  setBlock("D19", "BARCODE_ID", 26) %>%
  setBlock("D20", "BARCODE_ID", 27) %>%
  setBlock("D21", "BARCODE_ID", 28) %>%
  setBlock("K08", "BARCODE_ID", 22) %>%
  setBlock("K09", "BARCODE_ID", 23) %>%
  setBlock("K10", "BARCODE_ID", 24) %>%
  setBlock("K11", "BARCODE_ID", 25) %>%
  setBlock("K12", "BARCODE_ID", 26) %>%
  setBlock("K13", "BARCODE_ID", 27) %>%
  setBlock("K14", "BARCODE_ID", 28) %>%  
  setBlock("N08", "BARCODE_ID", 22) %>%
  setBlock("N09", "BARCODE_ID", 23) %>%
  setBlock("N10", "BARCODE_ID", 24) %>%
  setBlock("N11", "BARCODE_ID", 25) %>%
  setBlock("N12", "BARCODE_ID", 26) %>%
  setBlock("N13", "BARCODE_ID", 27) %>%
  setBlock("N14", "BARCODE_ID", 28) %>% 
  setBlock("E01", "BARCODE_ID", 29) %>%
  setBlock("E02", "BARCODE_ID", 30) %>%
  setBlock("E03", "BARCODE_ID", 31) %>%
  setBlock("E04", "BARCODE_ID", 32) %>%
  setBlock("E05", "BARCODE_ID", 33) %>%
  setBlock("E06", "BARCODE_ID", 34) %>%
  setBlock("E07", "BARCODE_ID", 35) %>%
  setBlock("E08", "BARCODE_ID", 29) %>%
  setBlock("E09", "BARCODE_ID", 30) %>%
  setBlock("E10", "BARCODE_ID", 31) %>%
  setBlock("E11", "BARCODE_ID", 32) %>%
  setBlock("E12", "BARCODE_ID", 33) %>%
  setBlock("E13", "BARCODE_ID", 34) %>%
  setBlock("E14", "BARCODE_ID", 35) %>%  
  setBlock("E15", "BARCODE_ID", 29) %>%
  setBlock("E16", "BARCODE_ID", 30) %>%
  setBlock("E17", "BARCODE_ID", 31) %>%
  setBlock("E18", "BARCODE_ID", 32) %>%
  setBlock("E19", "BARCODE_ID", 33) %>%
  setBlock("E20", "BARCODE_ID", 34) %>%
  setBlock("E21", "BARCODE_ID", 35) %>%  
  setBlock("L08", "BARCODE_ID", 29) %>%
  setBlock("L09", "BARCODE_ID", 30) %>%
  setBlock("L10", "BARCODE_ID", 31) %>%
  setBlock("L11", "BARCODE_ID", 32) %>%
  setBlock("L12", "BARCODE_ID", 33) %>%
  setBlock("L13", "BARCODE_ID", 34) %>%
  setBlock("L14", "BARCODE_ID", 35) %>%  
  setBlock("O08", "BARCODE_ID", 29) %>%
  setBlock("O09", "BARCODE_ID", 30) %>%
  setBlock("O10", "BARCODE_ID", 31) %>%
  setBlock("O11", "BARCODE_ID", 32) %>%
  setBlock("O12", "BARCODE_ID", 33) %>%
  setBlock("O13", "BARCODE_ID", 34) %>%
  setBlock("O14", "BARCODE_ID", 35) %>%    
  setBlock("F01", "BARCODE_ID", 36) %>%
  setBlock("F02", "BARCODE_ID", 37) %>%
  setBlock("F03", "BARCODE_ID", 38) %>%
  setBlock("F04", "BARCODE_ID", 39) %>%
  setBlock("F05", "BARCODE_ID", 40) %>%
  setBlock("F06", "BARCODE_ID", 41) %>%
  setBlock("F07", "BARCODE_ID", 42) %>% 
  setBlock("F08", "BARCODE_ID", 36) %>%
  setBlock("F09", "BARCODE_ID", 37) %>%
  setBlock("F10", "BARCODE_ID", 38) %>%
  setBlock("F11", "BARCODE_ID", 39) %>%
  setBlock("F12", "BARCODE_ID", 40) %>%
  setBlock("F13", "BARCODE_ID", 41) %>%
  setBlock("F14", "BARCODE_ID", 42) %>%  
  setBlock("F15", "BARCODE_ID", 36) %>%
  setBlock("F16", "BARCODE_ID", 37) %>%
  setBlock("F17", "BARCODE_ID", 38) %>%
  setBlock("F18", "BARCODE_ID", 39) %>%
  setBlock("F19", "BARCODE_ID", 40) %>%
  setBlock("F20", "BARCODE_ID", 41) %>%
  setBlock("F21", "BARCODE_ID", 42) %>%
  setBlock("M08", "BARCODE_ID", 36) %>%
  setBlock("M09", "BARCODE_ID", 37) %>%
  setBlock("M10", "BARCODE_ID", 38) %>%
  setBlock("M11", "BARCODE_ID", 39) %>%
  setBlock("M12", "BARCODE_ID", 40) %>%
  setBlock("M13", "BARCODE_ID", 41) %>%
  setBlock("M14", "BARCODE_ID", 42) %>%   
  setBlock("P08", "BARCODE_ID", 36) %>%
  setBlock("P09", "BARCODE_ID", 37) %>%
  setBlock("P10", "BARCODE_ID", 38) %>%
  setBlock("P11", "BARCODE_ID", 39) %>%
  setBlock("P12", "BARCODE_ID", 40) %>%
  setBlock("P13", "BARCODE_ID", 41) %>%
  setBlock("P14", "BARCODE_ID", 42) %>%  
  setBlock("G01", "BARCODE_ID", 43) %>%
  setBlock("G02", "BARCODE_ID", 44) %>%
  setBlock("G03", "BARCODE_ID", 45) %>%
  setBlock("G04", "BARCODE_ID", 46) %>%
  setBlock("G05", "BARCODE_ID", 47) %>%
  setBlock("G06", "BARCODE_ID", 48) %>%
  setBlock("G07", "BARCODE_ID", 49) %>% 
  setBlock("G08", "BARCODE_ID", 43) %>%
  setBlock("G09", "BARCODE_ID", 44) %>%
  setBlock("G10", "BARCODE_ID", 45) %>%
  setBlock("G11", "BARCODE_ID", 46) %>%
  setBlock("G12", "BARCODE_ID", 47) %>%
  setBlock("G13", "BARCODE_ID", 48) %>%
  setBlock("G14", "BARCODE_ID", 49) %>%   
  setBlock("G15", "BARCODE_ID", 43) %>%
  setBlock("G16", "BARCODE_ID", 44) %>%
  setBlock("G17", "BARCODE_ID", 45) %>%
  setBlock("G18", "BARCODE_ID", 46) %>%
  setBlock("G19", "BARCODE_ID", 47) %>%
  setBlock("G20", "BARCODE_ID", 48) %>%
  setBlock("G21", "BARCODE_ID", 49) %>% 
  setBlock("K15", "BARCODE_ID", 43) %>%
  setBlock("K16", "BARCODE_ID", 44) %>%
  setBlock("K17", "BARCODE_ID", 45) %>%
  setBlock("K18", "BARCODE_ID", 46) %>%
  setBlock("K19", "BARCODE_ID", 47) %>%
  setBlock("K20", "BARCODE_ID", 48) %>%
  setBlock("K21", "BARCODE_ID", 49) %>%   
  setBlock("N15", "BARCODE_ID", 43) %>%
  setBlock("N16", "BARCODE_ID", 44) %>%
  setBlock("N17", "BARCODE_ID", 45) %>%
  setBlock("N18", "BARCODE_ID", 46) %>%
  setBlock("N19", "BARCODE_ID", 47) %>%
  setBlock("N20", "BARCODE_ID", 48) %>%
  setBlock("N21", "BARCODE_ID", 49) %>%  
  setBlock("H01", "BARCODE_ID", 50) %>%
  setBlock("H02", "BARCODE_ID", 51) %>%
  setBlock("H03", "BARCODE_ID", 52) %>%
  setBlock("H04", "BARCODE_ID", 53) %>%
  setBlock("H05", "BARCODE_ID", 54) %>%
  setBlock("H06", "BARCODE_ID", 55) %>%
  setBlock("H07", "BARCODE_ID", 56) %>%
  setBlock("H08", "BARCODE_ID", 50) %>%
  setBlock("H09", "BARCODE_ID", 51) %>%
  setBlock("H10", "BARCODE_ID", 52) %>%
  setBlock("H11", "BARCODE_ID", 53) %>%
  setBlock("H12", "BARCODE_ID", 54) %>%
  setBlock("H13", "BARCODE_ID", 55) %>%
  setBlock("H14", "BARCODE_ID", 56) %>%
  setBlock("H15", "BARCODE_ID", 50) %>%
  setBlock("H16", "BARCODE_ID", 51) %>%
  setBlock("H17", "BARCODE_ID", 52) %>%
  setBlock("H18", "BARCODE_ID", 53) %>%
  setBlock("H19", "BARCODE_ID", 54) %>%
  setBlock("H20", "BARCODE_ID", 55) %>%
  setBlock("H21", "BARCODE_ID", 56) %>%
  setBlock("L15", "BARCODE_ID", 50) %>%
  setBlock("L16", "BARCODE_ID", 51) %>%
  setBlock("L17", "BARCODE_ID", 52) %>%
  setBlock("L18", "BARCODE_ID", 53) %>%
  setBlock("L19", "BARCODE_ID", 54) %>%
  setBlock("L20", "BARCODE_ID", 55) %>%
  setBlock("L21", "BARCODE_ID", 56) %>%  
  setBlock("O15", "BARCODE_ID", 50) %>%
  setBlock("O16", "BARCODE_ID", 51) %>%
  setBlock("O17", "BARCODE_ID", 52) %>%
  setBlock("O18", "BARCODE_ID", 53) %>%
  setBlock("O19", "BARCODE_ID", 54) %>%
  setBlock("O20", "BARCODE_ID", 55) %>%
  setBlock("O21", "BARCODE_ID", 56) %>% 
  setBlock("I01", "BARCODE_ID", 57) %>%
  setBlock("I02", "BARCODE_ID", 58) %>%
  setBlock("I03", "BARCODE_ID", 59) %>%
  setBlock("I04", "BARCODE_ID", 60) %>%
  setBlock("I05", "BARCODE_ID", 61) %>%
  setBlock("I06", "BARCODE_ID", 62) %>%
  setBlock("I07", "BARCODE_ID", 63) %>%  
  setBlock("I08", "BARCODE_ID", 57) %>%
  setBlock("I09", "BARCODE_ID", 58) %>%
  setBlock("I10", "BARCODE_ID", 59) %>%
  setBlock("I11", "BARCODE_ID", 60) %>%
  setBlock("I12", "BARCODE_ID", 61) %>%
  setBlock("I13", "BARCODE_ID", 62) %>%
  setBlock("I14", "BARCODE_ID", 63) %>%    
  setBlock("I15", "BARCODE_ID", 57) %>%
  setBlock("I16", "BARCODE_ID", 58) %>%
  setBlock("I17", "BARCODE_ID", 59) %>%
  setBlock("I18", "BARCODE_ID", 60) %>%
  setBlock("I19", "BARCODE_ID", 61) %>%
  setBlock("I20", "BARCODE_ID", 62) %>%
  setBlock("I21", "BARCODE_ID", 63) %>%   
  setBlock("M15", "BARCODE_ID", 57) %>%
  setBlock("M16", "BARCODE_ID", 58) %>%
  setBlock("M17", "BARCODE_ID", 59) %>%
  setBlock("M18", "BARCODE_ID", 60) %>%
  setBlock("M19", "BARCODE_ID", 61) %>%
  setBlock("M20", "BARCODE_ID", 62) %>%
  setBlock("M21", "BARCODE_ID", 63) %>%  
  setBlock("P15", "BARCODE_ID", 57) %>%
  setBlock("P16", "BARCODE_ID", 58) %>%
  setBlock("P17", "BARCODE_ID", 59) %>%
  setBlock("P18", "BARCODE_ID", 60) %>%
  setBlock("P19", "BARCODE_ID", 61) %>%
  setBlock("P20", "BARCODE_ID", 62) %>%
  setBlock("P21", "BARCODE_ID", 63) %>%   
  setBlock("J01", "BARCODE_ID", 64) %>%
  setBlock("J02", "BARCODE_ID", 65) %>%
  setBlock("J03", "BARCODE_ID", 66) %>%
  setBlock("J04", "BARCODE_ID", 67) %>%
  setBlock("J05", "BARCODE_ID", 68) %>%
  setBlock("J06", "BARCODE_ID", 69) %>%
  setBlock("J07", "BARCODE_ID", 70) %>%   
  setBlock("J08", "BARCODE_ID", 64) %>%
  setBlock("J09", "BARCODE_ID", 65) %>%
  setBlock("J10", "BARCODE_ID", 66) %>%
  setBlock("J11", "BARCODE_ID", 67) %>%
  setBlock("J12", "BARCODE_ID", 68) %>%
  setBlock("J13", "BARCODE_ID", 69) %>%
  setBlock("J14", "BARCODE_ID", 70) %>%   
  setBlock("J15", "BARCODE_ID", 64) %>%
  setBlock("J16", "BARCODE_ID", 65) %>%
  setBlock("J17", "BARCODE_ID", 66) %>%
  setBlock("J18", "BARCODE_ID", 67) %>%
  setBlock("J19", "BARCODE_ID", 68) %>%
  setBlock("J20", "BARCODE_ID", 69) %>%
  setBlock("J21", "BARCODE_ID", 70) %>%  
  setBlock("M23", "BARCODE_ID", 64) %>%
  setBlock("K22", "BARCODE_ID", 65) %>%
  setBlock("K23", "BARCODE_ID", 66) %>%
  setBlock("K24", "BARCODE_ID", 67) %>%
  setBlock("L22", "BARCODE_ID", 68) %>%
  setBlock("L23", "BARCODE_ID", 69) %>%
  setBlock("L24", "BARCODE_ID", 70) %>%  
  setBlock("P23", "BARCODE_ID", 64) %>%
  setBlock("N22", "BARCODE_ID", 65) %>%
  setBlock("N23", "BARCODE_ID", 66) %>%
  setBlock("N24", "BARCODE_ID", 67) %>%
  setBlock("O22", "BARCODE_ID", 68) %>%
  setBlock("O23", "BARCODE_ID", 69) %>%
  setBlock("O24", "BARCODE_ID", 70)   
```

RT_PRIMERS
===========

RT primers concentration
-------------------------

 - 0 uM (controls, + 0 nL) in Col 1, 8 and 15 + M23 + P23
 - 0.125 uM (+ 25 nL from source well K1 2.5 uM) in Col 2, 9 and 16 + K22 + N22
 - 0.25 uM (+ 25 nL from source well K2 5 uM) in Col 3, 10 and 17 + K23 + N23
 - 0.5 uM (+ 25 nL from source well K3 10 uM) in Col 4, 11 and 18 + K24 + N24
 - 1 uM (+ 25 nL from source well K4 20 uM) in Col 5, 12 and 19 + L22 + O22
 - 2 uM (+ 25 nL from source well K5 40 uM) in Col 6, 13 and 20 + L23 + O23
 - 4 uM (+ 25 nL from source well K6 80 uM) in Col 7, 14 and 21 + L24 + O24

```{r RT_PRIMERS}
plate %<>%
  setBlock("A01~P01",   "RT_PRIMERS", 0) %>%
  setBlock("A08~P08",   "RT_PRIMERS", 0) %>%
  setBlock("A15~P15",   "RT_PRIMERS", 0) %>%
  setBlock("M23",   "RT_PRIMERS", 0) %>%
  setBlock("P23",   "RT_PRIMERS", 0) %>%
  setBlock("A02~P02",   "RT_PRIMERS", 0.125) %>%
  setBlock("A09~P09",   "RT_PRIMERS", 0.125) %>%
  setBlock("A16~P16",   "RT_PRIMERS", 0.125) %>%
  setBlock("K22",   "RT_PRIMERS", 0.125) %>%
  setBlock("N22",   "RT_PRIMERS", 0.125) %>%
  setBlock("A03~P03",   "RT_PRIMERS", 0.25) %>%
  setBlock("A10~P10",   "RT_PRIMERS", 0.25) %>%
  setBlock("A17~P17",   "RT_PRIMERS", 0.25) %>%
  setBlock("K23",   "RT_PRIMERS", 0.25) %>%
  setBlock("N23",   "RT_PRIMERS", 0.25) %>%
  setBlock("A04~P04",   "RT_PRIMERS", 0.5) %>%
  setBlock("A11~P11",   "RT_PRIMERS", 0.5) %>%
  setBlock("A18~P18",   "RT_PRIMERS", 0.5) %>%
  setBlock("K24",   "RT_PRIMERS", 0.5) %>%
  setBlock("N24",   "RT_PRIMERS", 0.5) %>%
  setBlock("A05~P05",   "RT_PRIMERS", 1) %>%
  setBlock("A12~P12",   "RT_PRIMERS", 1) %>%
  setBlock("A19~P19",   "RT_PRIMERS", 1) %>%
  setBlock("L22",   "RT_PRIMERS", 1) %>%
  setBlock("O22",   "RT_PRIMERS", 1) %>%
  setBlock("A06~P06",   "RT_PRIMERS", 2) %>%
  setBlock("A13~P13",   "RT_PRIMERS", 2) %>%
  setBlock("A20~P20",   "RT_PRIMERS", 2) %>%
  setBlock("L23",   "RT_PRIMERS", 2) %>%
  setBlock("O23",   "RT_PRIMERS", 2) %>%
  setBlock("A07~P07",   "RT_PRIMERS", 4) %>%
  setBlock("A14~P14",   "RT_PRIMERS", 4) %>%
  setBlock("A21~P21",   "RT_PRIMERS", 4) %>%
  setBlock("L24",   "RT_PRIMERS", 4) %>%
  setBlock("O24",   "RT_PRIMERS", 4) 
```

RT primer volume 
------------------

25 nl of RT_PRIMERS added in each well, except the negative controls 

```{r RT_PRIMERS_vol}
plate %<>%
  setBlock("A01~P01",   "RT_PRIMERS_vol", 0) %>%
  setBlock("A08~P08",   "RT_PRIMERS_vol", 0) %>%
  setBlock("A15~P15",   "RT_PRIMERS_vol", 0) %>%
  setBlock("M23",   "RT_PRIMERS_vol", 0) %>%
  setBlock("P23",   "RT_PRIMERS_vol", 0) %>%
  setBlock("A02~P02",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A09~P09",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A16~P16",   "RT_PRIMERS_vol", 25) %>%
  setBlock("K22",   "RT_PRIMERS_vol", 25) %>%
  setBlock("N22",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A03~P03",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A10~P10",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A17~P17",   "RT_PRIMERS_vol", 25) %>%
  setBlock("K23",   "RT_PRIMERS_vol", 25) %>%
  setBlock("N23",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A04~P04",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A11~P11",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A18~P18",   "RT_PRIMERS_vol", 25) %>%
  setBlock("K24",   "RT_PRIMERS_vol", 25) %>%
  setBlock("N24",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A05~P05",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A12~P12",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A19~P19",   "RT_PRIMERS_vol", 25) %>%
  setBlock("L22",   "RT_PRIMERS_vol", 25) %>%
  setBlock("O22",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A06~P06",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A13~P13",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A20~P20",   "RT_PRIMERS_vol", 25) %>%
  setBlock("L23",   "RT_PRIMERS_vol", 25) %>%
  setBlock("O23",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A07~P07",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A14~P14",   "RT_PRIMERS_vol", 25) %>%
  setBlock("A21~P21",   "RT_PRIMERS_vol", 25) %>%
  setBlock("L24",   "RT_PRIMERS_vol", 25) %>%
  setBlock("O24",   "RT_PRIMERS_vol", 25) 
```

RNA
====

RNA concentration
------------------

 - 0 (controls, + 0 nL) in J01-J21 + K22-K24 + L22-L24 + M23 + N22-N24 + O22-O24 + P23
 - 100 ng (+ 25 nL from source well L1 4 ug/uL) in A01-I07
 - 10 ng (+ 25 nL from source well L2 400 ng/uL) in A08-I14
 - 1 ng (+ 25 nL from source well L3 40 ng/uL) in A15-I21
 - 100 pg (+ 25 nL from source well L4 4 ng/uL) in K01-M21
 - 10 pg (+ 25 nL from source well L5 0.4 ng/uL) in N01-P21

```{r RNA}
plate %<>%
  setBlock("J01~J21", "RNA", 0) %>%
  setBlock("K22~L24", "RNA", 0) %>%
  setBlock("M23", "RNA", 0) %>%
  setBlock("N22~O24", "RNA", 0) %>%
  setBlock("P23", "RNA", 0) %>%
  setBlock("A01~I07", "RNA", 100) %>%
  setBlock("A08~I14", "RNA", 10) %>%
  setBlock("A15~I21", "RNA", 1) %>%
  setBlock("K01~M21", "RNA", 0.1) %>%
  setBlock("N01~P21", "RNA", 0.01)
```

RNA volume
-----------

25 nL of RNA added to each well, except the negative controls

```{r RNA_vol}
plate %<>%
  setBlock("J01~J21", "RNA_vol", 0) %>%
  setBlock("K22~L24", "RNA_vol", 0) %>%
  setBlock("M23", "RNA_vol", 0) %>%
  setBlock("N22~O24", "RNA_vol", 0) %>%
  setBlock("P23", "RNA_vol", 0) %>%
  setBlock("A01~I07", "RNA_vol", 25) %>%
  setBlock("A08~I14", "RNA_vol", 25) %>%
  setBlock("A15~I21", "RNA_vol", 25) %>%
  setBlock("K01~M21", "RNA_vol", 25) %>%
  setBlock("N01~P21", "RNA_vol", 25)
```

H2O
====

H2O volume
-----------

0, 25, 50, 75 or 100 nL H2O added to complete RT reaction volume up to 500 nL

```{r H2O_vol}
plate %<>%
  setBlock("A02~A07", "H2O_vol", 0) %>%
  setBlock("A09~A14", "H2O_vol", 0) %>%
  setBlock("A16~A21", "H2O_vol", 0) %>%
  setBlock("D02~D07", "H2O_vol", 0) %>%
  setBlock("D09~D14", "H2O_vol", 0) %>%
  setBlock("D16~D21", "H2O_vol", 0) %>%
  setBlock("G02~G07", "H2O_vol", 0) %>%
  setBlock("G09~G14", "H2O_vol", 0) %>%
  setBlock("G16~G21", "H2O_vol", 0) %>%
  setBlock("K02~K07", "H2O_vol", 0) %>%
  setBlock("K09~K14", "H2O_vol", 0) %>%
  setBlock("K16~K21", "H2O_vol", 0) %>%
  setBlock("N02~N07", "H2O_vol", 0) %>%
  setBlock("N09~N14", "H2O_vol", 0) %>%
  setBlock("N16~N21", "H2O_vol", 0) %>%
  setBlock("J02~J07", "H2O_vol", 25) %>%
  setBlock("J09~J14", "H2O_vol", 25) %>%
  setBlock("J16~J21", "H2O_vol", 25) %>%
  setBlock("J02~J07", "H2O_vol", 25) %>%
  setBlock("K22~K24", "H2O_vol", 25) %>%
  setBlock("L22~L24", "H2O_vol", 25) %>%
  setBlock("N22~N24", "H2O_vol", 25) %>%
  setBlock("O22~O24", "H2O_vol", 25) %>%
  setBlock("K22~K24", "H2O_vol", 25) %>%
  setBlock("A01", "H2O_vol", 25) %>%
  setBlock("A08", "H2O_vol", 25) %>%
  setBlock("A15", "H2O_vol", 25) %>%
  setBlock("D01", "H2O_vol", 25) %>%
  setBlock("D08", "H2O_vol", 25) %>%
  setBlock("D15", "H2O_vol", 25) %>%
  setBlock("G01", "H2O_vol", 25) %>%
  setBlock("G08", "H2O_vol", 25) %>%
  setBlock("G15", "H2O_vol", 25) %>%
  setBlock("K01", "H2O_vol", 25) %>%
  setBlock("K08", "H2O_vol", 25) %>%
  setBlock("K15", "H2O_vol", 25) %>%
  setBlock("N01", "H2O_vol", 25) %>%
  setBlock("N08", "H2O_vol", 25) %>%
  setBlock("N15", "H2O_vol", 25) %>%
  setBlock("B02~B07", "H2O_vol", 50) %>%
  setBlock("B09~B14", "H2O_vol", 50) %>%
  setBlock("B16~B21", "H2O_vol", 50) %>%
  setBlock("E02~E07", "H2O_vol", 50) %>%
  setBlock("E09~E14", "H2O_vol", 50) %>%
  setBlock("E16~E21", "H2O_vol", 50) %>%
  setBlock("H02~H07", "H2O_vol", 50) %>%
  setBlock("H09~H14", "H2O_vol", 50) %>%
  setBlock("H16~H21", "H2O_vol", 50) %>%
  setBlock("L02~L07", "H2O_vol", 50) %>%
  setBlock("L09~L14", "H2O_vol", 50) %>%
  setBlock("L16~L21", "H2O_vol", 50) %>%
  setBlock("O02~O07", "H2O_vol", 50) %>%
  setBlock("O09~O14", "H2O_vol", 50) %>%
  setBlock("O16~O21", "H2O_vol", 50) %>%
  setBlock("M23", "H2O_vol", 50) %>%
  setBlock("P23", "H2O_vol", 50) %>%
  setBlock("J01", "H2O_vol", 50) %>%
  setBlock("J08", "H2O_vol", 50) %>%
  setBlock("J15", "H2O_vol", 50) %>%
  setBlock("C02~C07", "H2O_vol", 75) %>%
  setBlock("C09~C14", "H2O_vol", 75) %>%
  setBlock("C16~C21", "H2O_vol", 75) %>%
  setBlock("F02~F07", "H2O_vol", 75) %>%
  setBlock("F09~F14", "H2O_vol", 75) %>%
  setBlock("F16~F21", "H2O_vol", 75) %>%
  setBlock("I02~I07", "H2O_vol", 75) %>%
  setBlock("I09~I14", "H2O_vol", 75) %>%
  setBlock("I16~I21", "H2O_vol", 75) %>%
  setBlock("M02~M07", "H2O_vol", 75) %>%
  setBlock("M09~M14", "H2O_vol", 75) %>%
  setBlock("M16~M21", "H2O_vol", 75) %>%
  setBlock("P02~P07", "H2O_vol", 75) %>%
  setBlock("P09~P14", "H2O_vol", 75) %>%
  setBlock("P16~P21", "H2O_vol", 75) %>%
  setBlock("B01", "H2O_vol", 75) %>%
  setBlock("B08", "H2O_vol", 75) %>%
  setBlock("B15", "H2O_vol", 75) %>%
  setBlock("E01", "H2O_vol", 75) %>%
  setBlock("E08", "H2O_vol", 75) %>%
  setBlock("E15", "H2O_vol", 75) %>%
  setBlock("H01", "H2O_vol", 75) %>%
  setBlock("H08", "H2O_vol", 75) %>%
  setBlock("H15", "H2O_vol", 75) %>%
  setBlock("L01", "H2O_vol", 75) %>%
  setBlock("L08", "H2O_vol", 75) %>%
  setBlock("L15", "H2O_vol", 75) %>%
  setBlock("O01", "H2O_vol", 75) %>%
  setBlock("O08", "H2O_vol", 75) %>%
  setBlock("O15", "H2O_vol", 75) %>%
  setBlock("C01", "H2O_vol", 100) %>%
  setBlock("C08", "H2O_vol", 100) %>%
  setBlock("C15", "H2O_vol", 100) %>%
  setBlock("F01", "H2O_vol", 100) %>%
  setBlock("F08", "H2O_vol", 100) %>%
  setBlock("F15", "H2O_vol", 100) %>%
  setBlock("I01", "H2O_vol", 100) %>%
  setBlock("I08", "H2O_vol", 100) %>%
  setBlock("I15", "H2O_vol", 100) %>%
  setBlock("M01", "H2O_vol", 100) %>%
  setBlock("M08", "H2O_vol", 100) %>%
  setBlock("M15", "H2O_vol", 100) %>%
  setBlock("P01", "H2O_vol", 100) %>%
  setBlock("P08", "H2O_vol", 100) %>%
  setBlock("P15", "H2O_vol", 100)
```

RATIO TSO/RT_PRIMERS
=====================

Different ratio of TSO/RT_PRIMERS tested

```{r ratio_TSO_RT_PRIMERS}
plate$PRIMERS_RATIO <- ""
plate$PRIMERS_RATIO <- c(plate$TSO/plate$RT_PRIMERS)
plate$PRIMERS_RATIO <- sub("Inf", "no_RT_PRIMERS", plate$PRIMERS_RATIO)
```

```{r}
plate$total_volume <- ""
plate$total_volume <- rowSums(plate[, c(3,5,8,10,11)])
plate 
plate$total_volume
plate$BARCODE_ID
length(which(plate$total_volume != "NA")) 
sum(is.na(plate$total_volume))
count(plate$BARCODE_ID)
plate$PRIMERS_RATIO
```
