---
title: "Dummy version to make an explanatory figure"
output: 
  html_document: 
    fig_height: 5.25
    fig_width: 5.25
    keep_md: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 100)
knitr::opts_chunk$set(dev  = c("png", "svg"))
```

```{r}
variant <- "a"
newseed <- 1
indexes <- c("TAAGGCGA", "CGTACTAG", "AGGCAGAA", "TCCTGAGC", "GGACTCCT", "TAGGCATG")
plateID <- "Q"
```


Dummy version to make an explanatory figure
=================================================

This file starts as a copy of Labcyte-RT6.Rmd and may still
contain commands that are not really needed to make the figure panel,
but that I do not have time to remove.

## Load scripts and libraries.

```{r}
library("magrittr")   # For its diamond pipe operators
library("ggplot2")    # For plotting
library("platetools") # For defining and manipulating 384-well plates
library("plyr")
library("tibble")     # For comfortable command-line operations

# The set_block function that will become part of platetools
source("https://raw.githubusercontent.com/charles-plessy/platetools/76213accb1704d11a2d96fb1f6284d0b46117778/R/set_block.R")

# Do not load the whole smallCAGEqc package, just get the barcode data.
data("nanoCAGE2017barcodes", package = "smallCAGEqc")
```


Selected TSOs
=============

```{r tiers}
Tier_N <- 1:64
```

Creation of sextants
====================

```{r sextants, dependson=c("tiers", "setRandomSeed")}
createRandomisedSextant <- function(index, rna) {
    df <- data.frame ( BARCODE_ID =  Tier_N)
  tsoMaxConc <- 800
  df$TSO_source <- c( rep(tsoMaxConc /  1, 21)
                    , rep(tsoMaxConc /  8, 21)
                    , rep(tsoMaxConc /  64, 21)
                    , tsoMaxConc / 32)
  df$TSO_vol <- c(rep(c(rep(100,7), rep(50,7), rep(25,7)),3), 25)
  df$TSO <- df$TSO_source * df$TSO_vol / 500
  df$RT_PRIMERS <- c(rep(c(0, 1, 2, 4, 8, 16, 24), 9), 1)
  df$RT_PRIMERS_vol <- 25
  df[df$RT_PRIMERS == 0, "RT_PRIMERS_vol"] <- 0
  df$MASTER_MIX_vol <- 350
  df$INDEX <- index
  df$RNA <- rna
  df$RNA_vol <- 25
  df[64, "RNA_vol"] <- 0
  df$H2O_vol <- 500 - df$RNA_vol - df$RT_PRIMERS_vol - df$TSO_vol - df$MASTER_MIX_vol
  df
}
```


Sextants
========

Plate layout
------------

Create a data frame representing the contents of each well.

```{r make_plate_df}
plate           <- tibble(well = platetools::num_to_well(1:384, plate = "384"))
plate$row       <- sub("..$", "", plate$well) %>% factor
plate$col       <- sub("^.",  "", plate$well) %>% as.numeric %>% factor
```

Fill sextants
-------------

```{r fill_sextants, dependson=c("sextants", "make_plate_df")}
indexes <- c("TAAGGCGA", "CGTACTAG", "AGGCAGAA", "TCCTGAGC", "GGACTCCT", "TAGGCATG")

plate$sxt <- paste0 (
  c(rep("A", 8), rep("B", 8), rep("C", 8)),
  ifelse (plate$row %in% LETTERS[1:8 * 2], 2, 1)
) %>% factor

plate %<>% arrange(sxt)

plate %<>% cbind( rbind( createRandomisedSextant(indexes[1], 1e5)
                       , createRandomisedSextant(indexes[2], 1e4)
                       , createRandomisedSextant(indexes[3], 1e3)
                       , createRandomisedSextant(indexes[4], 1e2)
                       , createRandomisedSextant(indexes[5], 1e1)
                       , createRandomisedSextant(indexes[6], 1e0))
      ) %>% as.tibble()

plate$BARCODE_ID
plate$INDEX %<>% factor
plate$BARCODE_SEQ <- nanoCAGE2017barcodes[plate$BARCODE_ID, "barcodes"]

plate$RNA_level <- plate$RNA %>% factor(labels = paste0("RNA_", LETTERS[1:6]))
plate$RTP_level <- plate$RT_PRIMERS %>% factor(labels = paste0("RTP_", LETTERS[1:7]))

plate$RNA[plate$RNA_vol == 0] <- 0
plate$plateID <- "Z"
```


Plate maps
==========

```{r}
plateMap <- function(x, title) {
  platetools::raw_map(plate[[x]], well=plate$well, plate="384") +
  ggtitle(title) +
  viridis::scale_fill_viridis(breaks = unique(plate[[x]]))
}

plateMapLog <- function(x, title) {
  platetools::raw_map(plate[[x]], well=plate$well, plate="384") +
  ggtitle(title) +
  viridis::scale_fill_viridis(breaks = unique(plate[[x]]), trans = "log")
}
```

TSO
---

### Without legend, to better scale with the index plot.

```{r plot_barcode_IDs, fig.height=3}
platetools::raw_map(plate[["BARCODE_ID"]], well=plate$well, plate="384") +
  ggtitle("Barcode ID") + theme(legend.position="none") + viridis::scale_fill_viridis()
```

### With legend

```{r plot_barcode_IDs_legend, fig.height=3}
platetools::raw_map(plate[["BARCODE_ID"]], well=plate$well, plate="384") +
  ggtitle("Barcode ID") + viridis::scale_fill_viridis(breaks = c(1, 64))
```

Indexes
-------

### Without legend, to better scale with the barcode plot.

```{r plot_Indexes, fig.height=3}
platetools::raw_map(plate[["INDEX"]], well=plate$well, plate="384") + ggtitle("INDEXes") + theme(legend.position="none")
```

### With legend

```{r plot_Indexes_legend, fig.height=3}
platetools::raw_map(plate[["INDEX"]], well=plate$well, plate="384") + ggtitle("INDEXes")
```

## Session information

```{r sessionInfo}
sessionInfo()
```