---
title: "Concentration control of the TSO in the source plates of exp. 7"
output: 
  html_document: 
    fig_height: 5.25
    fig_width: 5.25
    keep_md: yes
    toc: yes
    pandoc_args:
     - "--lua-filter=links-to-markdown.lua"
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 100)
```


Load R packages
===============

```{r load_libraries}
library("magrittr")
library("ggplot2")
```


Load data
=========

Concentrations of the actual TSO solutions in source plates, measured with
a NanoDrop 8000, that has a dynamic range of 2.5–3,700 ng/μL for dsDNA according
to the manufacturer's [website](https://www.thermofisher.com/jp/en/home/industrial/spectroscopy-elemental-isotope-analysis/molecular-spectroscopy/ultraviolet-visible-visible-spectrophotometry-uv-vis-vis/uv-vis-vis-instruments/nanodrop-microvolume-spectrophotometers/nanodrop-products-guide.html).


Source plate
------------

Original file name `180510_kato.xlsx`.  This file is for the source plate used
in [experiment 7](Labcyte-RT_Data_Analysis_7.md) (transfer design 6).  See
[6a](Labcyte-RT6a.md) for example and expected concentrations.

```{r concentration_QC_load_2}
source <- gdata::read.xls( "Labcyte-Exp7.quantification.xlsx"
                         , nrow = 210, sheet = 1, stringsAsFactors = FALSE)
source$Plate.ID <- "source"

source$col <- source$Well. %>% sub(pat = ".", rep = "")
source$row <- source$Well. %>% substr(1,1)
source$orig_row <- source$col %>% factor(levels = 1:10, labels = c("A", "C", "E", "G", "I", "B", "D", "F", "H", "J"))
source$orig_col <- source$row %>% factor %>% as.numeric
source$orig_well <- paste0(source$orig_row, sprintf("%02d", source$orig_col))

conc <- data.frame( Well  = source$orig_well      
                  , plate = source$Plate.ID  %>% factor
                  , A260  = source$A260.
                  , A280  = source$A280.)

conc$Well %<>% factor(levels = levels(conc$Well) %>% gtools::mixedsort())
```


Load maker's information
------------------------

```{r idt_table}
idt <- read.csv("TSO_master_plate_PO_8268526.csv")
idt <- idt[,c("Well.Position", "Extinction.Coefficient.L..mole.cm.")]
conc$ext <- idt[match(conc$Well, idt$Well), "Extinction.Coefficient.L..mole.cm."]
```


Calculate concentrations (in micromolars)
-----------------------------------------

```{r calc_concentration, dependson=c("idt_table", "concentration_QC")}
conc$obs <- conc$A260 / conc$ext * 1e6
```


Average replicates
------------------

```{r concentration_average_replicates, dependson = "calc_concentration"}
conc.sd <- aggregate( conc[,c("obs", "A260", "A280")]
                    , list(Well = conc$Well, plate = conc$plate)
                    , sd)

conc <-    aggregate( conc[,c("obs", "A260", "A280")]
                    , list(Well = conc$Well, plate = conc$plate)
                    , mean)

summary(conc)
```


Expected concentrations
-----------------------

```{r exp_conc, dependson="concentration_average_replicates"}
conc[grepl("[ABC]", conc$Well),  "exp"] <- 800 / 50
conc[grepl("[DEFJ]", conc$Well), "exp"] <- 200 / 20
conc[grepl("[GHI]", conc$Well),  "exp"] <-  50 / 10
```

Histograms
==========

```{r concentration_QC_histograms, warning = FALSE, dependson = "exp_conc"}
hist_obs  <- ggplot(conc, aes(obs,  fill = exp %>% factor)) + geom_histogram() + ggtitle("Observed concentrations (ng/μL)")
hist_a260 <- ggplot(conc, aes(A260, fill = exp %>% factor)) + geom_histogram() + ggtitle("Absorbance at 260 nm")
hist_a280 <- ggplot(conc, aes(A280, fill = exp %>% factor)) + geom_histogram() + ggtitle("Absorbance at 280 nm")

ggpubr::ggarrange( ncol = 1, nrow = 3, hist_obs, hist_a260, hist_a280)
```


Absorbances and concentration
=============================

## A260 vs A280

```{r concentration_QC_abs_ratio}
ggplot(conc, aes(A280, A260, colour = exp)) + geom_point() +
    scale_x_log10() + scale_y_log10() +
  ggtitle("Relation between absorbances at 260 and 280 nm")

ggplot(conc[conc$A260 > 1,], aes(A280, A260, colour = exp)) + geom_point() +
  ggtitle("Relation between absorbances at 260 and 280 nm (outliers removed)")
```

## concentration vs A260 after the empty wells

```{r concentration_QC_conc_a260}
ggplot(conc, aes(obs, A260, colour = exp)) + geom_point() +
  ggtitle("Relation between concentration and absorbances at 260 nm")
```


Session information
===================

```{r sesion_info}
sessionInfo()
```