---
title: "TSO concentration in exp. 7's source plate"
output: 
  html_document: 
    fig_height: 5.25
    fig_width: 5.25
    keep_md: yes
    toc: yes
    pandoc_args:
     - "--lua-filter=links-to-markdown.lua"
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 100)
```

Concentrations of the actual TSO solutions in source plates, measured with
a NanoDrop 8000, that has a dynamic range of 2.5–3,700 ng/μL for dsDNA according
to the manufacturer's [website](https://www.thermofisher.com/jp/en/home/industrial/spectroscopy-elemental-isotope-analysis/molecular-spectroscopy/ultraviolet-visible-visible-spectrophotometry-uv-vis-vis/uv-vis-vis-instruments/nanodrop-microvolume-spectrophotometers/nanodrop-products-guide.html).


Load R packages
===============

```{r load_libraries}
library("magrittr")
library("ggplot2")
```


Load data
=========

Source plate
------------

Original file name `180510_kato.xlsx`.  This file is for the source plate used
in [experiment 7](Labcyte-RT_Data_Analysis_7.md), which uses transfer design 6.
See [design 6a](Labcyte-RT6a.md) for example and expected concentrations.


```{r concentration_QC_load_1}
source <- gdata::read.xls( "Labcyte-Exp7.quantification.xlsx"
                         , nrow = 210, sheet = 1, stringsAsFactors = FALSE)
source$Plate.ID <- "exp7"

source$col <- source$Well. %>% sub(pat = ".", rep = "")
source$row <- source$Well. %>% substr(1,1)
```

TSOs were transferred from the source plate to a dillution plate, in new
coordinates documented in sheet 2.

```{r concentration_QC_load_2, dependson="concentration_QC_load_1"}
source$orig_row  <- source$col %>%
                      factor( levels = 1:10,
                              labels = c("A", "C", "E", "G", "I", "B", "D", "F", "H", "J"))
source$orig_col  <- source$row %>% factor %>% as.numeric
source$orig_well <- paste0(source$orig_row, sprintf("%02d", source$orig_col))
```


Average replicates
------------------

```{r concentration_average_replicates, dependson = "concentration_QC_load_2"}
conc <- data.frame( Well  = source$orig_well
                  , plate = source$Plate.ID
                  , A260  = source$A260.
                  , A280  = source$A280.
                  , stringsAsFactors = FALSE)

conc.sd <- aggregate( conc[,c("A260", "A280")]
                    , list(Well = conc$Well, plate = conc$plate)
                    , sd)

conc <-    aggregate( conc[,c("A260", "A280")]
                    , list(Well = conc$Well, plate = conc$plate)
                    , mean)

summary(conc)
```


Barcode IDs
-----------

See [design 6a](Labcyte-RT6a.md) for example and expected concentrations if you
would like to double-check.

```{r}
conc$ID <- c(  3, 15, 27, 39, 51, 63, 75
            , 10, 22, 34, 46, 58, 70, 82
            , 11, 23, 35, 47, 59, 71, 83
            ,  1, 13, 25, 37, 49, 61, 73
            ,  2, 14, 26, 38, 50, 62, 74
            ,  7, 19, 31, 43, 55, 67, 79
            ,  4, 16, 28, 40, 52, 64, 76
            ,  5, 17, 29, 41, 65, 77, 89
            ,  6, 18, 30, 42, 66, 78, 90
            ,  8, 20, 32, 44, 68, 80, 92)

conc$stock_Well <- platetools::num_to_well(1:96)[conc$ID]
```


Load maker's information
------------------------

```{r idt_table}
idt <- read.csv("TSO_master_plate_PO_8268526.csv")
idt <- idt[,c("Well.Position", "Extinction.Coefficient.L..mole.cm.")]
conc$ext <- idt[match(conc$stock_Well, idt$Well), "Extinction.Coefficient.L..mole.cm."]
```


Calculate concentrations (in micromolars)
-----------------------------------------

```{r calc_concentration, dependson=c("idt_table", "concentration_QC")}
conc$obs <- conc$A260 / conc$ext * 1e6
```



Expected concentrations
-----------------------

To stay in the dynamic range, a different dilution factor was applied to the
TSOs according to their expected concentrations.

```{r exp_conc, dependson="concentration_average_replicates"}
conc[grepl("[ABC]",  conc$Well), "expected"] <- 800 / 50
conc[grepl("[DEFJ]", conc$Well), "expected"] <- 200 / 20
conc[grepl("[GHI]",  conc$Well), "expected"] <-  50 / 10
conc$dil <- 1000 / conc$expected
conc$expected %<>% paste(" µM") %>% factor
conc$expected %<>% factor(levels = levels(.) %>% gtools::mixedsort())
```


Histograms
==========

```{r concentration_QC_histograms, warning = FALSE, dependson = "exp_conc"}
hist_obs  <- ggplot(conc, aes(obs,  fill = expected)) + geom_histogram() + ggtitle("Observed concentrations (ng/μL)")
hist_a260 <- ggplot(conc, aes(A260, fill = expected)) + geom_histogram() + ggtitle("Absorbance at 260 nm")
hist_a280 <- ggplot(conc, aes(A280, fill = expected)) + geom_histogram() + ggtitle("Absorbance at 280 nm")

ggpubr::ggarrange( ncol = 1, nrow = 3, hist_obs, hist_a260, hist_a280)
```


Absorbances and concentration
=============================

## A260 vs A280

```{r concentration_QC_abs_ratio}
ggplot(conc, aes(A280, A260, colour = expected)) + geom_point() +
    scale_x_log10() + scale_y_log10() +
  ggtitle("Relation between absorbances at 260 and 280 nm")
```

## concentration vs A260

```{r concentration_QC_conc_a260}
ggplot(conc, aes(obs, A260, colour = expected)) + geom_point() +
  ggtitle("Relation between concentration and abs. at 260 nm")
```



Comparison between source and stock
===================================

The source plate was made without concentrations of the stock primers (see
[TSO_concentration_check2](TSO_concentration_check2.md) for details).

Here, we check whether the deviations of concentrations seen in the stock are
also obsereved in the source plate.

```{r dil_factors, fig.height=3, fig.width=9.5}
conc$stock <-
  read.table( "dilution_table.txt"
            , sep = "\t"
            , header = TRUE)[,"source_obs_molarity", drop = T][conc$ID]

conc$exp <- conc$stock / conc$dil

ggplot(conc, aes(obs, exp, colour = expected)) + geom_point() +
  facet_wrap(~expected, scale = "free")
```


Session information
===================

```{r sesion_info}
sessionInfo()
```