---
title: "TSO concentration in the stock plate PO_8268526"
output: 
  html_document: 
    fig_height: 5.25
    fig_width: 5.25
    keep_md: yes
    toc: yes
    pandoc_args:
     - "--lua-filter=links-to-markdown.lua"
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 100)
```

Concentrations of the actual TSO solutions in the stock plate PO_8268526, measured with
a NanoDrop 8000, that has a dynamic range of 2.5–3,700 ng/μL for dsDNA according
to the manufacturer's [website](https://www.thermofisher.com/jp/en/home/industrial/spectroscopy-elemental-isotope-analysis/molecular-spectroscopy/ultraviolet-visible-visible-spectrophotometry-uv-vis-vis/uv-vis-vis-instruments/nanodrop-microvolume-spectrophotometers/nanodrop-products-guide.html).

Load R packages
===============

```{r load_libraries}
library("magrittr")
library("ggplot2")
```


Load data
=========

Original file name `20180419_TSOs_stf.txt`.  This file has one sheet containing
raw data of old TSOs (stock plate PO_8268526, first IDT plate ordered) diluted
10× from the original source plate (all TSOs resuspended at 1 mM) and thus at an expected concentration of 100 μM.

```{r concentration_QC_load_IDs}
conc <- gdata::read.xls( "20180419_TSOs_stf2.xlsx"
                        , stringsAsFactors = FALSE)
```

```{r create_conc_table}
conc <- data.frame( Well  = conc$Well
                  , A260  = conc$A260.
                  , A280  = conc$A280.
                  , obs_conc = conc$Conc..
                  , stringsAsFactors = FALSE)

conc$col <- conc$Well %>% sub(pat = ".", rep = "") %>% as.integer
conc$row <- conc$Well %>% substr(1,1)
conc$Well <- paste0(conc$row, sprintf("%02d", conc$col)) %>% factor
conc$Well %<>% factor(levels = levels(conc$Well) %>% gtools::mixedsort())
```

Average replicates.

```{r concentration_QC_average_replicates, dependson = "concentration_QC_merge"}
conc <- aggregate( conc[,c("obs_conc", "A260", "A280")]
                 , list(Well = conc$Well)
                 , mean)
```

Load maker's information.
 
```{r idt_table}
idt <- read.csv("TSO_master_plate_PO_8268526.csv", dec = ",")
idt <- idt[,c("Well.Position", "OD260", "nmoles", "µg", "Measured.Molecular.Weight", "Extinction.Coefficient.L..mole.cm.")]
idt$ug.OD <- idt$µg / idt$OD260

conc$ext <- idt[match(conc$Well, idt$Well), "Extinction.Coefficient.L..mole.cm."]
conc$obs_source_MW <- idt[match(conc$Well, idt$Well), "Measured.Molecular.Weight"]
conc$conc_corr_Fact <- idt[match(conc$Well, idt$Well), "ug.OD"]
```


Calculate concentrations 
------------------------

### Molarity using extinction coefficients

```{r calc_concentration, dependson=c("idt_table", "concentration_QC")}
conc$obs <- conc$A260 / conc$ext * 1e6
```

### Mass using µg / OD

The nanoDrop reported mass using a coefficient of 50 µg per OD.  They are
corrected here using the information provided by the maker.

```{r correct_concentration, dependson=c("idt_table", "concentration_QC_load_molarity")}
conc$corr_obs_conc <- conc$obs_conc / 50 * conc$conc_corr_Fact
```

The expected molarity of TSOs in this measure is 100 μM.
The average molecular weight of nanoCAGE TSOs = ~14,000 g/mol, according to IDT plate report.
Therefore the expected concentration of TSOs at 100 μM (100 μmol/L) should be:
100 μmol/L * 14000 g/mol = 1.4 g/L = 1400 ng/μL.

```{r concentration_QC_load_molarity_and expected_concentration, dependson = "concentration_QC_average_replicates"}
conc$exp_molarity <- 100
conc$exp_conc <- conc$exp_molarity * conc$obs_source_MW / 1000
conc$obs_molarity <- (conc$corr_obs_conc*1000) / conc$obs_source_MW
summary(conc)
cor(conc$obs, conc$obs_molarity)
```

The valuse calcutated using calcultated extinction coefficients or using
empirical µg / OD factors fit well.


Find outliers.
--------------

```{r TSOs_to_discard_low_concentration}
conc$out_range <- "other"
conc[conc$obs_molarity < 25, "out_range"] <- "down"
conc[conc$out_range == "down",]
```

```{r TSOs_to_discard_high_concentration}
conc[conc$obs_molarity > 175, "out_range"] <- "up"
conc[conc$out_range == "up",]
```

Find TSOs in the range of expected concentrations.

```{r TSOs_in_concentration_range_expected_+/-_25%}
conc$delta_molarity <- conc$exp_molarity - conc$obs_molarity
conc[abs(conc$delta_molarity) < 25, "out_range"] <- "in"
conc[conc$out_range == "in", "Well"] %>% as.character
conc$out_range %<>% factor
```


Histograms
==========

```{r concentration_QC_histograms, warning = FALSE, dependson = "load_libraries", fig.height = 6.5, fig.width = 6.5}
hist_obs_molarity  <- ggplot(conc, aes(obs_molarity,  fill = out_range)) + geom_histogram() + ggtitle("Observed molarity (µM)")

hist_a260 <- ggplot(conc, aes(A260, fill = out_range)) + geom_histogram() +
  ggtitle("Absorbance at 260 nm")

hist_a280 <- ggplot(conc, aes(A280, fill = out_range)) + geom_histogram() +
  ggtitle("Absorbance at 280 nm")

ggpubr::ggarrange( ncol = 1, nrow = 4, hist_obs_molarity, hist_a260, hist_a280)
```



Absorbances and concentration
=============================

## A260 vs A280

```{r concentration_QC_abs_ratio}
ggplot(conc, aes(A280, A260, colour = exp_molarity)) + geom_point() +
    scale_x_log10() + scale_y_log10() +
  ggtitle("Relation between absorbances at 260 and 280 nm")
```

## concentration vs A260

```{r concentration_QC_obs_exp2}
ggplot(conc, aes(A260,  corr_obs_conc, colour = out_range)) + geom_point() +
  scale_x_log10() + scale_y_log10() +
  ggtitle("Observed concentration and expected concentration")
```


Dilution factors of TSOs
=========================

```{r concentrations_in_source_plate_and_dilution_factor, dependson = "concentration_QC_average_replicates"}
conc$source_obs_molarity <- conc$obs_molarity * 10
conc$dilution_factor_for_100uM <- conc$source_obs_molarity / 100
conc$TSO_vol_for_100uM <- 20 / conc$dilution_factor_for_100uM
conc$H20_vol_for_100uM <- 20 - conc$TSO_vol_for_100uM
conc$TSO_vol_for_100uM_2 <- 3
conc$H2O_vol_for_100uM_2 <- ((conc$TSO_vol_for_100uM_2 * conc$source_obs_molarity) / conc$exp_molarity) - conc$TSO_vol_for_100uM_2
```


Add visual comments
====================

```{r TSOs_to_discard}
do_not_use <- subset(conc, conc$dilution_factor_for_100uM < 2) 
do_not_use
nrow(do_not_use)
do_not_use$Well
```

Create dilution table.

```{r dilution_tables}
dilution_table <- conc[, c("Well", "source_obs_molarity", "dilution_factor_for_100uM", "TSO_vol_for_100uM", "H20_vol_for_100uM")]
```

Print tables.

```{r save_file}
write.table(dilution_table, "dilution_table.txt", row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
```


Session information
===================

```{r sesion_info}
sessionInfo()
```
