---
title: "Concentration control of the TSO stock and source plates of exp. 6"
output: 
  html_document: 
    fig_height: 5.25
    fig_width: 5.25
    keep_md: yes
    toc: yes
    pandoc_args:
     - "--lua-filter=links-to-markdown.lua"
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 100)
```


Load R packages
===============

```{r load_libraries}
library("magrittr")
library("ggplot2")
```


Load data
=========

Concentrations of the actual TSO solutions in source plates, measured with
a NanoDrop 8000, that has a dynamic range of 2.5–3,700 ng/μL for dsDNA according
to the manufacturer's [website](https://www.thermofisher.com/jp/en/home/industrial/spectroscopy-elemental-isotope-analysis/molecular-spectroscopy/ultraviolet-visible-visible-spectrophotometry-uv-vis-vis/uv-vis-vis-instruments/nanodrop-microvolume-spectrophotometers/nanodrop-products-guide.html).


Stock plate
-----------

Original file name `20180419_TSOs_stf.txt`.  This file has one sheet containing
raw data of old TSOs (first IDT plate ordered) diluted 10× from the original
stock plate (all TSOs resuspended at 1 mM) and thus at an expected concentration
of 100 μM.

```{r concentration_QC_load_1}
stock <- gdata::read.xls( "old_TSOs_conc/20180419_TSOs_stf2.xlsx"
                        , stringsAsFactors = FALSE)
stock$Plate.ID <- "stock"
stock$X <- NA  # For rbind() later
```


Source plate
------------

Original file name `180507_kato.xlsx`.  This file is for the source plate used
in [experiment 6](Labcyte-RT_Data_Analysis_6.md)
(transfer design [6](Labcyte-RT6.md)).  Expected concentration: 100 μM.
Samples have been dilluted 10 times before measuring concentration (see sheet 4).

```{r concentration_QC_load_2}
source <- gdata::read.xls( "Labcyte-RT6.quantification.xlsx"
                         , nrow = 288, sheet = 1, stringsAsFactors = FALSE)
source$Plate.ID <- "source"
source$A260. %<>% multiply_by(10)  # Dilution factor
source$A280. %<>% multiply_by(10)  # Dilution factor
```


Combine plates
--------------

```{r concentration_QC, dependson = c("concentration_QC_load_1", "concentration_QC_load_2")}
conc <- rbind(stock, source)

conc <- data.frame( Well  = conc$Well      
                  , plate = conc$Plate.ID  %>% factor
                  , A260  = conc$A260.
                  , A280  = conc$A280.)

conc$Well %<>% factor(levels = levels(conc$Well) %>% gtools::mixedsort())
conc$exp <- 100
```


Load maker's information
------------------------

```{r idt_table}
idt <- read.csv("old_TSOs_conc/TSO_master_plate_PO_8268526.csv")
idt <- idt[,c("Well.Position", "Extinction.Coefficient.L..mole.cm.")]
idt$Well <- idt$Well.Position %>% sub(pat = "0(.)", rep = "\\1")
conc$ext <- idt[match(conc$Well, idt$Well), "Extinction.Coefficient.L..mole.cm."]
```


Calculate concentrations (in micromolars)
-----------------------------------------

```{r calc_concentration, dependson=c("idt_table", "concentration_QC")}
conc$obs <- conc$A260 / conc$ext * 1e6
```


Average replicates
------------------

```{r concentration_average_replicates, dependson = "calc_concentration"}
conc.sd <- aggregate( conc[,c("obs", "exp", "A260", "A280")]
                    , list(Well = conc$Well, plate = conc$plate)
                    , sd)

conc <-    aggregate( conc[,c("obs", "exp", "A260", "A280")]
                    , list(Well = conc$Well, plate = conc$plate)
                    , mean)

summary(conc)
```


Histograms
==========

```{r concentration_QC_histograms, warning = FALSE, dependson = "concentration_average_replicates"}
hist_obs  <- ggplot(conc, aes(obs,  fill = plate)) + geom_histogram() +
  facet_wrap(~plate, nrow = 1) + ggtitle("Observed concentrations (ng/μL)")
hist_a260 <- ggplot(conc, aes(A260, fill = plate)) + geom_histogram() +
  facet_wrap(~plate, nrow = 1) + ggtitle("Absorbance at 260 nm")
hist_a280 <- ggplot(conc, aes(A280, fill = plate)) + geom_histogram() +
  facet_wrap(~plate, nrow = 1) + ggtitle("Absorbance at 280 nm")

ggpubr::ggarrange( ncol = 1, nrow = 3, hist_obs, hist_a260, hist_a280)
```


Absorbances and concentration
=============================

## A260 vs A280

```{r concentration_QC_abs_ratio}
ggplot(conc, aes(A280, A260, colour = plate)) + geom_point() +
    scale_x_log10() + scale_y_log10() +
  ggtitle("Relation between absorbances at 260 and 280 nm")

ggplot(conc[conc$A260 > 1,], aes(A280, A260, colour = plate)) + geom_point() +
  ggtitle("Relation between absorbances at 260 and 280 nm (outliers removed)")
```

## concentration vs A260 after the empty wells

```{r concentration_QC_conc_a260}
ggplot(conc[conc$obs > 10,], aes(obs, A260, colour = plate)) + geom_point() +
  ggtitle("Relation between concentration and absorbances at 260 nm")
```


Comparison between source and stock
===================================

The source plate was made by adjusting concentrations of the stock primers,
based on the measurement of their concentration, see
[old_TSOs_conc/TSO_conc2](old_TSOs_conc/TSO_conc2.md) for details.

Here, we verify that the dilution factors were applied as expected.  

```{r dil_factors, dependson = "correct_concentration"}
dil <- data.frame( Well = levels(conc$Well)
                 , exp  = read.table("old_TSOs_conc/dilution_table2.txt", dec = ",", sep = "\t", header = TRUE)[,2, drop = T]
                 , obs  = conc[conc$plate == "stock", "obs"] / conc[conc$plate == "source", "obs"])

with(dil[dil$obs > 0 & dil$ obs < 100,], plot(obs, exp))
```


Session information
===================

```{r sesion_info}
sessionInfo()
```