---
title: "Xmas"
author: "Charles Plessy"
date: "26 December 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load scripts and libraries.

```{r}
library("magrittr")   # For its diamond pipe operators
library("ggplot2")    # For plotting
library("platetools") # For defining and manipulating 384-well plates
library("plyr")
library("tibble")     # For comfortable command-line operations

# The set_block function that will become part of platetools
source("https://raw.githubusercontent.com/charles-plessy/platetools/76213accb1704d11a2d96fb1f6284d0b46117778/R/set_block.R")
```

## Plate layout

Create a data frame representing the contents of each well.

```{r make_plate}
plate <- tibble(well = platetools::num_to_well(1:384, plate = "384"))
plate
```

```{r add_stem}
plate %<>% set_block("H21~I24", "stem", 1)
```

```{r add_leaves}
plate %<>%
  set_block("B19~O20", "leaves", 1) %>%
  set_block("C17~N18", "leaves", 1) %>%
  set_block("D15~M16", "leaves", 1) %>%
  set_block("E13~L14", "leaves", 1) %>%
  set_block("F09~K12", "leaves", 1) %>%
  set_block("G05~J08", "leaves", 1) %>%
  set_block("H02~I04", "leaves", 1)
```

```{r add_Xmas_balls}
plate %<>%
  set_block("G05", "balls", 1) %>%
  set_block("J08", "balls", 1) %>%
  set_block("F13", "balls", 1) %>%
  set_block("J12", "balls", 1) %>%
  set_block("I17", "balls", 1) %>%
  set_block("E18", "balls", 1) %>%
  set_block("M16", "balls", 1)
```

```{r plot_plate}
plateMap <- function(plate, x, title) {
  platetools::raw_map(plate[[x]], well=plate$well, plate="384") +
  ggtitle(title) +
  viridis::scale_fill_viridis(breaks = unique(plate[[x]]))
}

(plot_stem   <- plateMap(plate, "stem",   "stem"))
(plot_leaves <- plateMap(plate, "leaves", "leaves"))
(plot_balls  <- plateMap(plate, "balls",  "balls"))


x <- data.frame(lapply(colnames(plate[,-1]), function(x) ifelse(is.na(plate[,x]), "", x)))
x <- apply( x, 1
          , function(x)
              if(all(x=="")) {
                NA
              } else {
                paste(unique(x[x!=""]), collapse=",")
              })
plate$contents <- x

raw_map(plate$contents, well=plate$well, plate="384")
```

## Source plate

```{r}
colSums(plate[,-1], na.rm = TRUE)

source <- tibble(well = platetools::num_to_well(1:384, plate = "384"))
source %<>%
  set_block("A01", "stem", 20) %>%
  set_block("A02~A04", "leaves", 60) %>%
  set_block("A05", "balls", 20)

x <- data.frame(lapply(colnames(source[,-1]), function(x) ifelse(is.na(source[,x]), "", x)))
x <- apply( x, 1
          , function(x)
              if(all(x=="")) {
                NA
              } else {
                paste(unique(x[x!=""]), collapse=",")
              })
source$contents <- x

raw_map(source$contents, well=source$well, plate="384")
```

## Transfer system

Let `l` be a list comprising

 - `source` (a source plate)
 - `plan` (a plate representing the final goal)
 - `target` (a target plate)
 - `transfers` (a list of transfer commands)
 
The function `transfer` will take `l` as input and return a list where

 - `source` is the source plate, with volumes removed
 - `plan` is unchanged
 - `target` is the target plates with volumes added
 - `transfers` is the list of transfer commands, with new transfers appended.
 
(Later, a vectorised version may be written with a list of source plates as input)
 

```{r}


transfer <- function(l) {
  source <- l$source
  
  
  ...
  
  return(l)
}

```